### 2014-07-09 paper/paper.tex
Course paper outline

\documentclass{sig-alternate-2013}

\pdfminorversion=4
\pdfimageresolution=600

\usepackage{caption}
\usepackage{subfig}
\usepackage{todonotes}
\usepackage[hidelinks]{hyperref}
\usepackage{fixltx2e}
\usepackage{enumitem}
\usepackage[english]{babel}
\usepackage{microtype}

\clubpenalty=10000
\widowpenalty=10000

\makeatletter
\def\@copyrightspace{\relax}
\makeatother

\begin{document}

\hypersetup{pdftitle=Congestion-Aware Join Optimization in Key-Value Stores}
\hypersetup{pdfauthor={Xu Cui, Michael Mior, and Xinan Yan}}

\title{Congestion-Aware Join Optimization in Key-Value Stores}

\numberofauthors{3}
\author{
    \alignauthor{
        Xu Cui\\
        \affaddr{University of Waterloo}\\
        \email{xcui@uwaterloo.ca}
    }
    \alignauthor{
        Michael Mior\\
        \affaddr{University of Waterloo}\\
        \email{mmior@uwaterloo.ca}
    }
    \alignauthor{
        Xinan Yan\\
        \affaddr{University of Waterloo}\\
        \email{x29yan@uwaterloo.ca}
    }
}

\maketitle

\begin{abstract}
\end{abstract}

\section{Introduction}

\section{System Architecture}

\section{Related Work}

\cite{Xiong2014}

\section{Evaluation}

\section{Future Work}

\section{Conclusion}

\let\theOLDbibliography\thebibliography\renewcommand{\thebibliography}[1]{\theOLDbibliography{#1}%
\item[]\vspace*{0.5mm}}

\bibliographystyle{abbrv}
{\scriptsize
\bibliography{paper}
}

\end{document}

### 2014-07-15 paper/paper.tex
Start of some paper sections

@@
 \begin{abstract}
+Applications which use key-value stores to execute higher-level queries may need to follow references obtained from previous keys.
+In the case where all data is contained in main memory, the bottleneck becomes the network transfer of values from the datastore to the application.
+When some keys become highly popular, network links to machines hosting these keys can become saturated and affect application performance.
+We demonstrate an approach that takes advantage of replication and network performance measurements to effectively route requests over underutilized links.
 \end{abstract}
@@
 
-\cite{Xiong2014}
+The majority of work on query processing in key-value stores is in executing queries on data stored in distributed hash tables (DHTs) in peer-to-peer (P2P) networks.
+For example, Harren et al.~\cite{Harren2002} propose a basic query model for executing complex queries in this scenario.
+Their approach aims to minimize communication cost, but it is designed for P2P systems where the network is not under control of the application and is therefore unable to make use of network-aware techniques.
+PIER~\cite{Huebsch2005} is a similar query processing engine over DHTs which is also not network-aware.
+However, the query processing model used by PIER may be an interesting area to explore.
 
@@
 
+We run all our experiments on a simulated network on a single machine using Mininet~\cite{Lantz2010}.
+
 \section{Future Work}
+
+There are many opportunities to combine this work with appropriate key placement strategies for further performance optimizations.
+For heterogenous networks, a key placement strategy which is network and workload aware may provide significant performance advantages.
+Our current query execution model is very simplistic and could be expanded to support richer queries and distributed index data structures.
+We would also like to investigate other query types such as aggregation which we suspect would also benefit from our approach.
+Finally, our current system requires all intermediate values to be sent to the application which then coordinates all requests.
+Making the datastore a participant in query execution would reduce network bandwidth by allowing individual servers to follow the links between keys.
 

### 2014-07-30 paper/paper.tex
Additional paper text

@@
 
-\section{System Architecture}
+\section{System Overview}
+
+\subsection{Query Model}
+
+\subsection{Architecture}
+
+If we were to use fair share allocation, then the node with the largest share of the query results, bottlenecked by available bandwidth, would become a straggler.
+Nodes with a small share of query results could finish much more quickly only to wait for these stragglers.
+To minimize the overall completion time of queries, we can allocate bandwidth to each node as the fraction of query results which will be returned by the node.
+This will speed up stragglers at the expense of slowing down nodes with a smaller allocation.
+However, the net result is a reduction in the average execution cost.
+
+We currently deal with a read-only workload where the distribution of data is known by all nodes.
+Each node is able to calculate the fraction of query results it will return.
 
@@
 
+We also have not dealt with the possibility of updates as well as replication.
+There is potential to further exploit SDN techniques by allowing low latency updates to data in the presence of the high bandwidth flows used for query processing.
+
 \section{Conclusion}

### 2014-07-30 paper/paper.tex
More paper text

@@
 
+\subsection{Flow Scheduling}
+
+Simple approaches to bandwith allocation such as max-min fairness pose significant problems in our settings.
 If we were to use fair share allocation, then the node with the largest share of the query results, bottlenecked by available bandwidth, would become a straggler.
@@
 Each node is able to calculate the fraction of query results it will return.
+A node can signal the expected fraction of the results for a query it will return by setting bits within the packet header.
+When a new flow is issued for query processing, the controller can inspect these bits and assign an appropriate priority to each flow.
 
@@
 We run all our experiments on a simulated network on a single machine using Mininet~\cite{Lantz2010}.
+We use Open vSwitch~\cite{Pfaff2009} as our switch which supports queues with varying bandwidth guarantees.
+Our controller uses the POX~\cite{Gude2008} platform and OpenFlow~\cite{McKeown2008} 1.0.
 
@@
 We would also like to investigate other query types such as aggregation which we suspect would also benefit from our approach.
-Finally, our current system requires all intermediate values to be sent to the application which then coordinates all requests.
-Making the datastore a participant in query execution would reduce network bandwidth by allowing individual servers to follow the links between keys.
+This presents new challenges in network scheduling, especially in the face of concurrent queries of multiple different types.
 
-We also have not dealt with the possibility of updates as well as replication.
+Our current system also does not deal with the possibility of updates or replication.
 There is potential to further exploit SDN techniques by allowing low latency updates to data in the presence of the high bandwidth flows used for query processing.
+Replication provides further opportunity for optimization as flows can be routed to multiple possible locations to avoid bottleneck links.
 

### 2014-07-31 src/CustomTopo.py
Committing Server & Client version 1

@@ -1,29 +1,35 @@
-'''
-Creating a data-center tree topology with link customization.
-For example fanout of 2 will generate
-             c1
-       a1            a2
-  e1      e2     e3      e4
-h1   h2 h3  h4 h5  h6  h7  h8
-where cX aX eX are switches and hX are hosts
-'''
-
 from mininet.topo import Topo
 from mininet.net import Mininet
 from mininet.log import setLogLevel
 from mininet.node import CPULimitedHost
 from mininet.link import TCLink
+from mininet.cli import CLI
+import os
+import sys
+import inspect
 
-class CustomTopo(Topo):
+# adds the current dir i.e. src to system path to include sshd
+sys.path.append(os.path.dirname(os.path.abspath(inspect.getfile(inspect.currentframe()))))
+from sshd import *
+
+class DataCenter(Topo):
     "Simple Data Center Topology"
 
+    """
+    Creating a data-center tree topology with link customization.
+    For example fanout of 2 will generate
+                  c1
+           a1            a2
+      e1      e2     e3      e4
+    h1   h2 h3  h4 h5  h6  h7  h8
+    where cX aX eX are switches and hX are hosts
+    """
     "linkopts - (1:core, 2:aggregation, 3: edge) parameters"
     "fanout - number of child switch per parent switch"
     def __init__(self, linkopts1, linkopts2, linkopts3, fanout=2, **opts):
         # Initialize topology and default options
         Topo.__init__(self, **opts)
 
-        # Add your logic here ...
         coreSwitch = self.addSwitch('c1')
         for i in range(fanout):
             aggregationSwitch = self.addSwitch('a' + str((i+1)))
@@ -36,19 +42,73 @@ class CustomTopo(Topo):
                     self.addLink(edgeSwitch, host, **linkopts3)
 
 
+class SimpleDC(Topo):
+    def __init__(self, linkopts1, linkopts2, fanout=2, **opts):
+        # Initialize topology and default options
+        Topo.__init__(self, **opts)
+        coreSwitch = self.addSwitch('c1')
+        for i in range(fanout):
+            aggregationSwitch = self.addSwitch('a' + str((i+1)))
[…]

### 2014-07-31 src/client.py

@@ -1,0 +1,33 @@
+# Echo client program
+import socket
+import json
+from test import *
+import time
+import sys
+
+NUM_HOST = 4
+ID_LIST = []
+
+def current_time_in_milli():
+    return int(round(time.time() * 1000))
+
+def loadIDs():
+    533721
+    pass
+
+# all_data = json.dumps(store.data[1])
+# sys.stdout.write(str(len(all_data)) + '\n')
+
+HOST = '10.0.0.4'    # The remote host
+PORT = 50007              # The same port as used by the server
+s = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+s.connect((HOST, PORT))
+
+all_data = 'abcd'
+start_time = current_time_in_milli()
+rtn = s.send(all_data)
+print rtn
+end_time = current_time_in_milli()
+print 'Total time spent to send the data %d' % (end_time - start_time)
+s.close()
+# print 'Received', len(data)

### 2014-07-31 src/scripts/clean.sh

@@ -1,0 +1,16 @@
+#!/bin/bash
+
+PROJECT_DIR='~/xcui/new_project/kvtest/src/'
+
+for i in h{1..4}
+do
+    ssh ${i} "kill -9 $(ps aux | grep server.p\[y\] | awk '{print $2}')"
+    echo server stopped at ${i}
+done
+
+
+for i in h{1..4}
+do
+    ssh ${i} "kill -9 $(ps aux | grep server.p\[y\] | awk '{print $2}')"
+    echo server stopped at ${i}
+done

### 2014-07-31 src/scripts/test.sh

@@ -1,0 +1,10 @@
+#!/bin/bash
+
+PROJECT_DIR='/media/sf_vm_shared/latest//new_project/kvtest/src/'
+
+rm -rf  ~/.ssh/known_hosts
+for i in h{1..4}
+do
+    ssh ${i} "cd ${PROJECT_DIR}; python ./server.py" &
+    echo server started at ${i}
+done

### 2014-07-31 src/server.py

@@ -1,0 +1,45 @@
+# Echo server program
+import socket
+import thread
+import sys
+import commands
+
+from storage import *
+
+
+NUM_HOST = 4
+HOST = ''    # Symbolic name meaning the local host
+# TODO: Read this port number form a configuration file
+PORT = 50007 # Arbitrary non-privileged port
+
+def getServerId():
+    # hostname = [(s.connect(('8.8.8.8', 80)), s.getsockname()[0], s.close()) for s in [socket.socket(socket.AF_INET, socket.SOCK_DGRAM)]][0][1]]
+    ip = commands.getoutput("/sbin/ifconfig").split("\n")[1].split()[1][5:]
+    rtn = int(ip.split('.')[3]) - 1
+    return rtn
+
+def serve(clientsocket, addr):
+    recv_size = 0
+    while 1:
+        data = clientsocket.recv(1024)
+        recv_size += len(data)
+        if not data: break
+    print >> sys.stderr, 'clientsocket %s is done and total data received = %s' % (str(addr), str(recv_size))
+    clientsocket.close()
+
+def startServer(sId = None, rFactor = 1):
+    store = Store(NUM_HOST, sId, rFactor)
+    users = User(store).read()
+    items = Item(store).read()
+    serverSocket = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+    serverSocket.bind((HOST, PORT))
+    serverSocket.listen(5)
+    while 1:
+        conn, addr = serverSocket.accept()
+        print >> sys.stderr, 'Connected by', addr
+        thread.start_new_thread(serve, (conn, addr))
+
+if __name__ == '__main__':
+    sId = getServerId()
+    print(sId)
+    startServer(sId, rFactor = 1)

### 2014-07-31 src/sshd.py

@@ -24,16 +24,21 @@ from mininet.log import lg
 from mininet.node import Node
 from mininet.topolib import TreeTopo
 from mininet.link import Link
-
-from CustomTopo import CustomTopo
 from mininet.node import CPULimitedHost
 from mininet.link import TCLink
+import os
+import inspect
+
+# adds the current dir i.e. src to system path to include sshd
+sys.path.append(os.path.dirname(os.path.abspath(inspect.getfile(inspect.currentframe()))))
+from CustomTopo import *
+
 
 def TreeNet( depth=1, fanout=2, **kwargs):
     "Convenience function for creating tree networks."
     # topo = TreeTopo( depth, fanout )
     topoLinkOpts = dict(bw=10, delay='5ms', loss=1, max_queue_size=1000, use_htb=True)
-    topo = CustomTopo(topoLinkOpts, topoLinkOpts, topoLinkOpts, fanout)
+    topo = DataCenter(topoLinkOpts, topoLinkOpts, topoLinkOpts, fanout)
     if not kwargs:
         kwargs = dict(host=CPULimitedHost, link=TCLink)
     return Mininet( topo, **kwargs )
@@ -86,5 +91,5 @@ if __name__ == '__main__':
     # useDNS=no -u0 to avoid reverse DNS lookup timeout
     opts = ' '.join( sys.argv[ 1: ] ) if len( sys.argv ) > 1 else (
         '-D -o UseDNS=no -u0' )
-    # our CustomTopo uses c1 as its root switch
+    # our DataCenter uses c1 as its root switch
     sshd( net, switch=net['c1'], opts=opts )

### 2014-07-31 src/storage.py

@@ -4,24 +4,11 @@ import os
 import sys
 import json
 
-# Add mininet to the path
-# sys.path.append(os.path.join(os.path.dirname(os.path.realpath(__file__)),
-                # 'mininet'))
-
 from mininet.topo import Topo
 
-CSV_BASE = 'xcui_csv/'
-
-class StarTopo(Topo):
-    def __init__(self, count):
-        self.count = count
-        Topo.__init__(self)
-
-        hosts = ['h%d' % i for i in range(count)]
-        switch = self.addSwitch('s')
-        for host in hosts:
-            self.addHost(host)
-            self.addLink(host, switch)
+# xcui's test csv
+# CSV_BASE = 'xcui_csv/'
+CSV_BASE = 'csv/'
 
 class Entity(object):
     foreign_keys = {}
@@ -43,17 +30,19 @@ class Entity(object):
 
             for row in reader:
                 id = row[0]
+                rowVal = dict()
                 for index, value in enumerate(row[1:]):
                     field = header[index + 1]
                     # Append the ID if we have a many-to-many relation or
                     # set the value directly otherwise
                     related = self.__class__.foreign_keys.get(field, None)
                     if related:
-                        self.store.append('%s:%s:%s' %
-                                (self.__class__.__name__, field, id), value)
+                        if field not in row:
+                            rowVal[field] = []
+                        rowVal[field].append(value)
                     else:
-                        self.store.set('%s:%s:%s' %
-                                (self.__class__.__name__, field, id), value)
+                        rowVal[field] = value
+                self.store.set(id, rowVal)
             # print(json.dumps(self.store.data))
 
 class User(Entity):
@@ -83,10 +72,10 @@ class Store(object):
     """
     def bucket(self, key):
         return hash(key) % len(self.data)
+
     """
     Set the value for the given key
     """
-
     def set(self, key, value):
         curBucket = self.bucket(key)
         if self.sId is not None:
@@ -124,7 +113,7 @@ In the current implementation, Query only works when we join all
 of the sharded tables together
 """
 class Query(object):
-    def __init__(self, entity, store):
+    def __init__(self, entity, store, IDs):
         self.entity = entity
         self.store = store
 
@@ -132,7 +121,7 @@ class Query(object):
     Execute a query to fetch the entity with the given ID and all its
     fields and related entities
     """
-    def execute(self, id):
[…]

### 2014-07-31 src/scripts/test.sh
Add Vagrant configuration

@@ -1,6 +1,6 @@
 #!/bin/bash
 
-PROJECT_DIR='/media/sf_vm_shared/latest//new_project/kvtest/src/'
+PROJECT_DIR='/mininet/src/'
 
 rm -rf  ~/.ssh/known_hosts
 for i in h{1..4}

### 2014-07-31 src/scripts/clean.sh
fixed errors in the cleanup script

@@ -1,16 +1,3 @@
 #!/bin/bash
 
-PROJECT_DIR='~/xcui/new_project/kvtest/src/'
-
-for i in h{1..4}
-do
-    ssh ${i} "kill -9 $(ps aux | grep server.p\[y\] | awk '{print $2}')"
-    echo server stopped at ${i}
-done
-
-
-for i in h{1..4}
-do
-    ssh ${i} "kill -9 $(ps aux | grep server.p\[y\] | awk '{print $2}')"
-    echo server stopped at ${i}
-done
+ps aux | grep server.py | awk '{print $2}' | xargs kill -9

### 2014-07-31 src/CustomTopo.py
Simplify Python path updates

@@ -6,10 +6,9 @@ from mininet.link import TCLink
 from mininet.cli import CLI
 import os
 import sys
-import inspect
 
 # adds the current dir i.e. src to system path to include sshd
-sys.path.append(os.path.dirname(os.path.abspath(inspect.getfile(inspect.currentframe()))))
+sys.path.append(os.path.dirname(__file__))
 from sshd import *
 
 class DataCenter(Topo):

### 2014-07-31 src/sshd.py

@@ -27,10 +27,9 @@ from mininet.link import Link
 from mininet.node import CPULimitedHost
 from mininet.link import TCLink
 import os
-import inspect
 
 # adds the current dir i.e. src to system path to include sshd
-sys.path.append(os.path.dirname(os.path.abspath(inspect.getfile(inspect.currentframe()))))
+sys.path.append(os.path.dirname(__file__))
 from CustomTopo import *
 
 

### 2014-08-01 paper/paper.tex
Expand on related work

@@
 
+There has also been some work in exploiting software-defined networking for big data applications.
+Wang et al.~\cite{Wang2012} examined network traffic patterns in a Hadoop cluster and identified opportunities to combine network-aware job scheduling with dynamic routing configuration to improve job completion time.
+Xiong et al.~\cite{Xiong2014} examine the usefulness of software-defined networking in supporting distributed analytical queries.
+Their workload consits of read-only SQL queries where query processing is bandwidth-intensive.
+They construct a global query optimizer which decides on join order and how query results should be passed between sites.
+This work takes a similar approach to ours, but also identifies many opportunities for further work.
+
 \section{Evaluation}

### 2014-08-01 paper/paper.tex
Add small note about RUBiS
Update paper abstract

@@
 \begin{abstract}
-Applications which use key-value stores to execute higher-level queries may need to follow references obtained from previous keys.
+Applications which use key-value stores to execute higher-level queries need to aggregate data from multiple nodes.
 In the case where all data is contained in main memory, the bottleneck becomes the network transfer of values from the datastore to the application.
 When some keys become highly popular, network links to machines hosting these keys can become saturated and affect application performance.
+In addition, nodes will compete with each other for bandwidth and uninformed traffic prioritization can result in reduced query execution time.
 We demonstrate an approach that takes advantage of replication and network performance measurements to effectively route requests over underutilized links.
@@
 
+Queries use data taken from the RUBiS~\cite{Cecchet2002} benchmark, which is a Web application simulating an online auction.
+We distribute data on users and the items they are selling between all nodes with no replication.
+
 \section{Future Work}

### 2014-08-01 paper/paper.tex
First pass at paper intro

@@
 In addition, nodes will compete with each other for bandwidth and uninformed traffic prioritization can result in reduced query execution time.
-We demonstrate an approach that takes advantage of replication and network performance measurements to effectively route requests over underutilized links.
+We demonstrate an approach based on software-defined networking (SDN) that takes advantage of replication and network performance measurements to effectively route requests over underutilized links.
 \end{abstract}
@@
 \section{Introduction}
+
+Recent trends have shown a proliferation of databases termed ``NoSQL'' which frequently eschew traditional database concepts such as referential integrity and ACID transactions for the sake of performance.
+However, applications using these databases still have a need to perform complex queries.
+Especially in the case of analytical queries, this may force the application designer to implement constructs such as joins at the application level.
+This requires a significant amount of data transfer between nodes.
+Given that it is common for data to be held in memory for performance reasons, the bottleneck for this data transfer is pushed up to the network layer.
+
+Poor network scheduling can easily result in suboptimal query execution.
+For example, consider the simple topology shown in XXXX.
+Assume a query is being processed that requires data to be sent between hosts A and C and hosts B and D that together will use the full capacity of all links in the network.
+If the flows have equal bandwidth requirements, then fair bandwidth allocation will perform well.
+However, if the flow between A and C requires twice the bandwidth of the flow between B and D, the first flow will have a longer completion time while it waits for available bandwidth.
+
+In this scenario, if we schedule traffic using weighted fair queuing and reserve additional bandwidth for the first flow, it will complete sooner.
+This results in a longer completion time for the second flow, but if we assume the query depends on all flows to be completed, we can see an overall reduction in query time.
+Similarly, in the face of replicated data, a network-aware query processing engine can select routes and replicas to maximize link utilization and improve performance.
+This type of intelligent network behaviour is difficult to configure in a traditional network stack.
+However, with the advent of SDN, it is possible for a smart SDN controller to make global decisions based on the knowledge of higher-level application queries.
 

### 2014-08-01 paper/paper.tex
Add dummy topology figure
Add start of query model section

@@
 
+\begin{figure}
+    \centering
+    \missingfigure{Topology graph}
+    \caption{Sample network topology}\label{fig:sample-topology}
+\end{figure}
+
 Poor network scheduling can easily result in suboptimal query execution.
-For example, consider the simple topology shown in XXXX.
+For example, consider the simple topology shown in Figure~\ref{fig:sample-topology}.
 Assume a query is being processed that requires data to be sent between hosts A and C and hosts B and D that together will use the full capacity of all links in the network.
@@
 \subsection{Query Model}
+
+\begin{figure}
+    \centering
+    \missingfigure{ER diagram}
+    \caption{Simple query entities}\label{fig:entities}
+\end{figure}
+
+Our prototype system considers foreign key join queries between entity sets with a fixed join order and simple filtering on attributes of the first entity set.
+For example, assume we have two entity sets \texttt{User} and \texttt{Item} with a one-to-many foreign key relating a user to each item as in Figure~\ref{fig:entities}.
+A query (in SQL) syntax may look like \texttt{SELECT * FROM User, Item WHERE Item.User\_id = User.id AND Item.price > 20}.
+In the future, we plan to consider a richer query language.
+This will also present more interesting opportunities for optimization.
+
+The execution model for these queries is very straightforward.
+We consider the execution of the simple query given above.
+We start by passing the query to each node which has entities of the class on the left side of the join (\texttt{Item}).
+Those nodes will scan their local list of entities and apply the necessary filters (\texttt{price > 20}).
+Each node then forwards the necessary foreign keys to nodes holding entites of the class on the right side of the join (\texttt{User}).
+These nodes are then able to look up entities on the other side of the join and results can be passed back to the client.
 

### 2014-08-01 src/joinHelpers.py
to merge test.sh

@@ -1,0 +1,15 @@
+import threading
+import threading
+
+class BarrierSync(object):
+    def __init__(self):
+        self.thread_count = 0
+        self.semaphore = threading.Semaphore()
+        self.event = threading.Event()
+    def sync(self, count):
+        """ All calls to this METHOD WILL block until the last (count) call is made """
+        with self.semaphore:
+            self.thread_count += 1
+            if self.thread_count == count:
+                self.event.set()
+        self.event.wait()

### 2014-08-01 src/scripts/test.sh

@@ -3,7 +3,7 @@
 PROJECT_DIR='/mininet/src/'
 
 rm -rf  ~/.ssh/known_hosts
-for i in h{1..4}
+for i in 10.0.0.{1..4}
 do
     ssh ${i} "cd ${PROJECT_DIR}; python ./server.py" &
     echo server started at ${i}

### 2014-08-01 src/client.py
server client initial commit

@@ -1,33 +1,101 @@
-# Echo client program
+#!/usr/bin/env python
 import socket
 import json
-from test import *
 import time
 import sys
+import csv
+import os
+import thread
+import threading
+import time
+from random import randint
+import inspect
+import traceback
+import struct
 
-NUM_HOST = 4
+# adds the current dir i.e. src to system path
+sys.path.append(os.path.dirname(os.path.abspath(inspect.getfile(inspect.currentframe()))))
+from joinHelpers import BarrierSync
+from joinHelpers import config
+from joinHelpers import err
+from joinHelpers import socket_read_n
+from joinHelpers import current_time_in_milli
+
+NUM_HOST = config['NUM_HOST']
 ID_LIST = []
+CSV_BASE = 'csv/'
 
-def current_time_in_milli():
-    return int(round(time.time() * 1000))
+BARRIER = BarrierSync()
 
-def loadIDs():
-    533721
-    pass
+SEMAPHORE = threading.Semaphore()
+FINAL_JOIN_RESULT = []
 
-# all_data = json.dumps(store.data[1])
-# sys.stdout.write(str(len(all_data)) + '\n')
+def loadItemIDs():
+    with open(CSV_BASE + 'items.csv') as csvfile:
+        reader = csv.reader(csvfile, delimiter=',', quotechar='\'',
+                            doublequote=False, escapechar='\\')
+        header = next(reader, None)
+        for row in reader:
+            id = row[0]
+            ID_LIST.append(int(id))
+    print ID_LIST
 
-HOST = '10.0.0.4'    # The remote host
-PORT = 50007              # The same port as used by the server
-s = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
-s.connect((HOST, PORT))
+def itemsJoin(host, port = 50007):
+    """ The actual function that performs the join operation on item & user """
+    # send
+    s = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+    s.connect((host, port))
+    allData = {config['OP']: config['JOIN_ITEM_USER_OP'], config['IDs']: ID_LIST}
+    toSend = json.dumps(allData)
+    # using big-endian to pack the data
+    dataLen = struct.pack('>L', len(toSend))
+    rtn = s.sendall(dataLen)
+    assert rtn == None
+    rtn = s.sendall(toSend)
+    assert rtn == None
+    lenBuf = s.recv(4)
[…]

### 2014-08-01 src/joinHelpers.py

@@ -1,5 +1,41 @@
 import threading
 import threading
+import time
+import sys
+
+config = {
+    'NUM_HOST': 4,  # Update when when change topology
+    'OP': 'op',
+    'ITEMS_OP': 'item',
+    'USERS_OP' : 'user',
+    # item is the primary key and user is the foreign key
+    'JOIN_ITEM_USER_OP' : 'join_item_user',
+    'IDs': 'ids',
+}
+
+def socket_read_n(sock, n):
+    """
+    Read exactly n bytes from the socket.
+    Raise RuntimeError if the connection closed before
+    n bytes were read.
+    """
+    buf = ''
+    while n > 0:
+        data = sock.recv(n)
+        if data == '':
+            raise RuntimeError('unexpected connection close')
+        buf += data
+        n -= len(data)
+    return buf
+
+def err(listOfInput):
+    sys.stderr.write('DEBUG\n')
+    if type(listOfInput) is list or type(listOfInput) is tuple:
+        for input in listOfInput:
+            sys.stderr.write(str(input) + ' ')
+    else:
+        sys.stderr.write(str(listOfInput))
+    sys.stderr.write('\nDEBUG\n')
 
 class BarrierSync(object):
     def __init__(self):
@@ -13,3 +49,6 @@ class BarrierSync(object):
             if self.thread_count == count:
                 self.event.set()
         self.event.wait()
+
+def current_time_in_milli():
+    return int(round(time.time() * 1000))

### 2014-08-01 src/scripts/test.sh

@@ -2,9 +2,10 @@
 
 PROJECT_DIR='/mininet/src/'
 
+mkdir -p ${PROJECT_DIR}/out
 rm -rf  ~/.ssh/known_hosts
 for i in 10.0.0.{1..4}
 do
-    ssh ${i} "cd ${PROJECT_DIR}; python ./server.py" &
+    ssh ${i} "cd ${PROJECT_DIR}; python ./server.py > out/server${i} 2>&1" &
     echo server started at ${i}
 done

### 2014-08-01 src/server.py

@@ -3,29 +3,98 @@ import socket
 import thread
 import sys
 import commands
+import json
+import threading
+import os
+import inspect
+import traceback
+import ast
+import struct
 
+# adds the current dir i.e. src to system path
+sys.path.append(os.path.dirname(os.path.abspath(inspect.getfile(inspect.currentframe()))))
+from joinHelpers import BarrierSync
+from joinHelpers import config
+from joinHelpers import err
+from joinHelpers import socket_read_n
 from storage import *
 
-
-NUM_HOST = 4
+NUM_HOST = int(config['NUM_HOST'])
 HOST = ''    # Symbolic name meaning the local host
 # TODO: Read this port number form a configuration file
 PORT = 50007 # Arbitrary non-privileged port
 
 def getServerId():
     # hostname = [(s.connect(('8.8.8.8', 80)), s.getsockname()[0], s.close()) for s in [socket.socket(socket.AF_INET, socket.SOCK_DGRAM)]][0][1]]
+    # print(commands.getoutput("/sbin/ifconfig").split("\n")[1])
     ip = commands.getoutput("/sbin/ifconfig").split("\n")[1].split()[1][5:]
     rtn = int(ip.split('.')[3]) - 1
     return rtn
 
-def serve(clientsocket, addr):
-    recv_size = 0
-    while 1:
-        data = clientsocket.recv(1024)
-        recv_size += len(data)
-        if not data: break
-    print >> sys.stderr, 'clientsocket %s is done and total data received = %s' % (str(addr), str(recv_size))
+def fetchUsers(sId, tuples, barrier):
+    pass
+
+def handleJoinItemUserOp(clientsocket, addr, storage, serverId, itemIds):
+    result = dict()
+    otherTraffic = {}
+    for id in itemIds:
+        itemId = 'items:' + str(id)
+        # There is no replication at the moment
+        itemServerId = storage.getServerIds(itemId)[0]
+        if(itemServerId == serverId):
+            itemRow = storage.get(itemId)
+            clientIds = itemRow['seller']
+            for cId in clientIds:
+                clientId = 'users:' + str(cId)
+                # Using [0] because we do not have replication at the moment
+                clientServerId = storage.getServerIds(clientId)[0]
+                if(str(serverId) == str(clientServerId)):
+                    itemRow['userInfo'] = storage.get(clientId)
+                    result[id] = itemRow
+                else:
[…]

### 2014-08-01 src/storage.py

@@ -1,3 +1,4 @@
+#!/usr/bin/env python
 import collections
 import csv
 import os
@@ -6,9 +7,11 @@ import json
 
 from mininet.topo import Topo
 
+from joinHelpers import err
+
 # xcui's test csv
-# CSV_BASE = 'xcui_csv/'
-CSV_BASE = 'csv/'
+CSV_BASE = 'xcui_csv/'
+# CSV_BASE = 'csv/'
 
 class Entity(object):
     foreign_keys = {}
@@ -29,7 +32,8 @@ class Entity(object):
             self.__class__.field_names = header
 
             for row in reader:
-                id = row[0]
+                id = str(row[0])
+                id = self.csv_name + ':' + id
                 rowVal = dict()
                 for index, value in enumerate(row[1:]):
                     field = header[index + 1]
@@ -63,7 +67,9 @@ class Store(object):
     """
     def __init__(self, num, sId = None, rFactor = 1):
         self.sId = sId
-        self.data = [collections.defaultdict(lambda: []) for _ in range(num)]
+        self.data = []
+        for i in range(num):
+            self.data.append({})
         if sId is not None:
             self.shard = [_ % num for _  in range(sId, sId + rFactor)]
 
@@ -74,6 +80,13 @@ class Store(object):
         return hash(key) % len(self.data)
 
     """
+    Return the sIds for which have this key
+    """
+    def getServerIds(self, key):
+        # We do not have replication at this stage, so we just return the hash
+        return [self.bucket(key)]
+
+    """
     Set the value for the given key
     """
     def set(self, key, value):
@@ -81,6 +94,9 @@ class Store(object):
         if self.sId is not None:
             if curBucket in self.shard:
                 self.data[curBucket][key] = value
+            else:
+                # not part this store's bucket
+                pass
         else:
             self.data[curBucket][key] = value
 
@@ -100,7 +116,8 @@ class Store(object):
     """
     def get(self, key):
         bucket = self.bucket(key)
-        return (bucket, self.data[bucket].get(key))
+        # return (bucket, self.data[bucket])
+        return self.data[bucket][key]
 
     """
     Return the number of keys stored in each bucket
@@ -135,37 +152,10 @@ class Query(object):
 if __name__ == '__main__':
     # Read a couple entities, run some queries with known IDs
     # and print the distribution of data between buckets
[…]

### 2014-08-01 paper/paper.tex
Add more system overview text

@@
 
+Our system uses OpenFlow~\cite{McKeown2008} enabled switches to allow centralized network control.
+We currently pass all queries through our SDN controller which is running on the POX~\cite{Gude2008} platform.
+
 \subsection{Flow Scheduling}
@@
 When a new flow is issued for query processing, the controller can inspect these bits and assign an appropriate priority to each flow.
+The controller is similar to a simple L2 learning switch.
+However, we configure several queues for each switch port which allow us to provide bandwidth reservations to support variation in expected traffic volume.
+
+We also change the flow table rules we install to match on the type of service (ToS) bits in the IP header.
+Two of these eight bits are often used for explicit congestion notification (ECN).
+We avoid these bits to allow ECN to operate uninhibited.
+The remaining six bits are the differentiated services code point (DSCP) which are used to provide varying QoS guarantees for different flows.
+We assume that the portion of the network our system is operating on is not carrying other traffic so we are free to set these bits as we choose.
+To interconnect with other networks which may assign different meaning to these bits, we can simply zero these bits for any traffic passing through ingress/egress switches.
 

### 2014-08-01 src/scripts/queues.sh
Add script for creating queues on OVS switches

@@ -1,0 +1,23 @@
+#!/bin/bash
+
+# Add queues with varying minimum bandwidth reservations to each switch port
+# queue 0 traffic will have priority over all other traffic
+
+for switch in $(sudo ovs-vsctl list-br); do
+    for port in $(sudo ovs-vsctl list-ports $switch); do
+        sudo ovs-vsctl -- set Port $port qos=@newqos -- \
+          --id=@newqos create QoS type=linux-htb other-config:max-rate=1000000000 \
+          queues=0=@q0,1=@q1,2=@q2,3=@q3,4=@q4,5=@q5,6=@q6,7=@q7,8=@q8,9=@q9,10=@q10 -- \
+          --id=@q0 create Queue other-config:priority=0 -- \
+          --id=@q1 create Queue other-config:min-rate=100000000 other-config:priority=1 -- \
+          --id=@q2 create Queue other-config:min-rate=200000000 other-config:priority=1 -- \
+          --id=@q3 create Queue other-config:min-rate=300000000 other-config:priority=1 -- \
+          --id=@q4 create Queue other-config:min-rate=400000000 other-config:priority=1 -- \
+          --id=@q5 create Queue other-config:min-rate=500000000 other-config:priority=1 -- \
+          --id=@q6 create Queue other-config:min-rate=600000000 other-config:priority=1 -- \
+          --id=@q7 create Queue other-config:min-rate=700000000 other-config:priority=1 -- \
+          --id=@q8 create Queue other-config:min-rate=800000000 other-config:priority=1 -- \
+          --id=@q9 create Queue other-config:min-rate=900000000 other-config:priority=1 -- \
+          --id=@q10 create Queue other-config:min-rate=1000000000 other-config:priority=1
+    done
+done

### 2014-08-03 src/client.py
full implementation of Server and Client

@@ -20,10 +20,12 @@ from joinHelpers import config
 from joinHelpers import err
 from joinHelpers import socket_read_n
 from joinHelpers import current_time_in_milli
+from joinHelpers import socket_send_data_by_json
+from joinHelpers import socket_recv_data_by_json
 
 NUM_HOST = config['NUM_HOST']
 ID_LIST = []
-CSV_BASE = 'csv/'
+CSV_BASE = config['CSV_BASE']
 
 BARRIER = BarrierSync()
 
@@ -31,14 +33,14 @@ SEMAPHORE = threading.Semaphore()
 FINAL_JOIN_RESULT = []
 
 def loadItemIDs():
-    with open(CSV_BASE + 'items.csv') as csvfile:
+    with open(CSV_BASE + '/items.csv') as csvfile:
         reader = csv.reader(csvfile, delimiter=',', quotechar='\'',
                             doublequote=False, escapechar='\\')
         header = next(reader, None)
         for row in reader:
             id = row[0]
             ID_LIST.append(int(id))
-    print ID_LIST
+    # print ID_LIST
 
 def itemsJoin(host, port = 50007):
     """ The actual function that performs the join operation on item & user """
@@ -46,21 +48,8 @@ def itemsJoin(host, port = 50007):
     s = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
     s.connect((host, port))
     allData = {config['OP']: config['JOIN_ITEM_USER_OP'], config['IDs']: ID_LIST}
-    toSend = json.dumps(allData)
-    # using big-endian to pack the data
-    dataLen = struct.pack('>L', len(toSend))
-    rtn = s.sendall(dataLen)
-    assert rtn == None
-    rtn = s.sendall(toSend)
-    assert rtn == None
-    lenBuf = s.recv(4)
-    dataLen = struct.unpack('>L', lenBuf)[0]
-    if(dataLen == 0):
-        partialResult = {}
-    else:
-        receivedRawData = socket_read_n(s, dataLen)
-        # err((receivedRawData, len(receivedRawData)))
-        partialResult = json.loads(receivedRawData)
+    socket_send_data_by_json(s, allData)
+    partialResult = socket_recv_data_by_json(s)
     with SEMAPHORE:
         FINAL_JOIN_RESULT.append(partialResult)
     s.close()
@@ -70,7 +59,6 @@ def joinThread(threadName, host):
     # time.sleep(randint(1,10))
     # Block until NUM_HOST + 1 threads have reached this point
[…]

### 2014-08-03 src/joinHelpers.py

@@ -2,9 +2,13 @@ import threading
 import threading
 import time
 import sys
+import json
+import struct
 
 config = {
+    'SERVER_PORT' : 50007,
     'NUM_HOST': 4,  # Update when when change topology
+    'CSV_BASE': 'csv/',  # source of RUBiS data
     'OP': 'op',
     'ITEMS_OP': 'item',
     'USERS_OP' : 'user',
@@ -12,6 +16,33 @@ config = {
     'JOIN_ITEM_USER_OP' : 'join_item_user',
     'IDs': 'ids',
 }
+
+def socket_send_data_by_json(sock, data):
+    toSend = json.dumps(data)
+    # using big-endian to pack the data
+    dataLen = struct.pack('>L', len(toSend))
+    rtn = sock.sendall(dataLen)
+    assert rtn == None
+    rtn = sock.sendall(toSend)
+    assert rtn == None
+    return
+
+def socket_recv_data_by_json(sock):
+    """
+    If data the msg_len == 0
+    this function returns None
+    otherwise this returns the data
+    """
+    len_buf = sock.recv(4)
+    # using big-endian to pack the data
+    msg_len = struct.unpack('>L', len_buf)[0]
+    if(msg_len == 0):
+        return None
+    else:
+        raw_data = socket_read_n(sock, msg_len)
+        assert len(raw_data) == msg_len
+        data = json.loads(raw_data)
+        return data
 
 def socket_read_n(sock, n):
     """

### 2014-08-03 src/scripts/clean.sh

@@ -1,3 +1,7 @@
 #!/bin/bash
 
-ps aux | grep server.py | awk '{print $2}' | xargs kill -9
+# kill servers
+ps aux | grep \.\/server\.py | awk '{print $2}' | xargs kill -9
+
+# kill the clients
+ps aux | grep \.\/client.py | awk '{print $2}' | xargs kill -9

### 2014-08-03 src/scripts/test.sh

@@ -6,6 +6,6 @@ mkdir -p ${PROJECT_DIR}/out
 rm -rf  ~/.ssh/known_hosts
 for i in 10.0.0.{1..4}
 do
-    ssh ${i} "cd ${PROJECT_DIR}; python ./server.py > out/server${i} 2>&1" &
+    ssh ${i} "cd ${PROJECT_DIR}; python ./server.py >> out/server${i} 2>&1" &
     echo server started at ${i}
 done

### 2014-08-03 src/server.py

@@ -17,12 +17,16 @@ from joinHelpers import BarrierSync
 from joinHelpers import config
 from joinHelpers import err
 from joinHelpers import socket_read_n
+from joinHelpers import socket_send_data_by_json
+from joinHelpers import socket_recv_data_by_json
 from storage import *
+
+
 
 NUM_HOST = int(config['NUM_HOST'])
 HOST = ''    # Symbolic name meaning the local host
 # TODO: Read this port number form a configuration file
-PORT = 50007 # Arbitrary non-privileged port
+PORT = int(config['SERVER_PORT']) # Arbitrary non-privileged port
 
 def getServerId():
     # hostname = [(s.connect(('8.8.8.8', 80)), s.getsockname()[0], s.close()) for s in [socket.socket(socket.AF_INET, socket.SOCK_DGRAM)]][0][1]]
@@ -31,69 +35,98 @@ def getServerId():
     rtn = int(ip.split('.')[3]) - 1
     return rtn
 
-def fetchUsers(sId, tuples, barrier):
-    pass
+def fetch_users(sId, tuples, result_container, result_sema, barrier,
+                waitNum, partial_result):
+    server_address = '10.0.0.' + str(sId + 1)
+    # this socket is talking to another server to fetch relevant user data
+    sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+    sock.connect((server_address, PORT))
+    allData = {config['OP'] : config['USERS_OP'],
+               config['IDs'] : tuples}
+    socket_send_data_by_json(sock, allData)
+    partial_seller_result = socket_recv_data_by_json(sock)
+    for key in partial_result:
+        # err(partial_seller_result)
+        # There is no replication at the moment so we can safely grab first item in [0]
+        seller_id = 'users:' + partial_result[key]['seller'][0]
+        seller_info = partial_seller_result[seller_id]
+        partial_result[key]['seller_info'] = seller_info
+    with result_sema:
+        result_container.update(partial_result)
+    sock.close()
+    barrier.sync(waitNum)
 
-def handleJoinItemUserOp(clientsocket, addr, storage, serverId, itemIds):
+def serve_user_op(client_socket, storage, client_ids):
+    rtn = dict()
+    for client_id in client_ids:
+        rtn[client_id] = storage.get(client_id)
+    socket_send_data_by_json(client_socket, rtn)
+
+def handleJoinItemUserOp(clientsocket, addr, storage, serverId, item_ids):
[…]

### 2014-08-03 src/storage.py

@@ -4,13 +4,19 @@ import csv
 import os
 import sys
 import json
+import inspect
 
 from mininet.topo import Topo
 
+
+# adds the current dir i.e. src to system path
+sys.path.append(os.path.dirname(os.path.abspath(inspect.getfile(inspect.currentframe()))))
 from joinHelpers import err
+from joinHelpers import config
+
 
 # xcui's test csv
-CSV_BASE = 'xcui_csv/'
+CSV_BASE = config['CSV_BASE']
 # CSV_BASE = 'csv/'
 
 class Entity(object):
@@ -48,6 +54,7 @@ class Entity(object):
                         rowVal[field] = value
                 self.store.set(id, rowVal)
             # print(json.dumps(self.store.data))
+        print >> sys.stderr, 'Loading Completed for %s...' % self.csv_name
 
 class User(Entity):
     csv_name = 'users'

### 2014-08-04 src/client.py
Simplify Python path updates

@@ -9,12 +9,11 @@ import thread
 import threading
 import time
 from random import randint
-import inspect
 import traceback
 import struct
 
 # adds the current dir i.e. src to system path
-sys.path.append(os.path.dirname(os.path.abspath(inspect.getfile(inspect.currentframe()))))
+sys.path.append(os.path.dirname(__file__))
 from joinHelpers import BarrierSync
 from joinHelpers import config
 from joinHelpers import err

### 2014-08-04 src/server.py

@@ -6,13 +6,12 @@ import commands
 import json
 import threading
 import os
-import inspect
 import traceback
 import ast
 import struct
 
 # adds the current dir i.e. src to system path
-sys.path.append(os.path.dirname(os.path.abspath(inspect.getfile(inspect.currentframe()))))
+sys.path.append(os.path.dirname(__file__))
 from joinHelpers import BarrierSync
 from joinHelpers import config
 from joinHelpers import err

### 2014-08-04 src/storage.py

@@ -4,13 +4,12 @@ import csv
 import os
 import sys
 import json
-import inspect
 
 from mininet.topo import Topo
 
 
 # adds the current dir i.e. src to system path
-sys.path.append(os.path.dirname(os.path.abspath(inspect.getfile(inspect.currentframe()))))
+sys.path.append(os.path.dirname(__file__))
 from joinHelpers import err
 from joinHelpers import config
 

### 2014-08-04 paper/paper.tex
Add missing diagrams for paper
Add more to architecture section

@@
     \centering
-    \missingfigure{Topology graph}
+    \includegraphics{figures/sample-topology.pdf}
     \caption{Sample network topology}\label{fig:sample-topology}
@@
     \centering
-    \missingfigure{ER diagram}
+    \includegraphics{figures/rubis-er.pdf}
     \caption{Simple query entities}\label{fig:entities}
@@
 We currently pass all queries through our SDN controller which is running on the POX~\cite{Gude2008} platform.
+Each host on the network contains a simple document store which is capable of answering requests both from client queries and other servers in the network.
+When a query is received from a client, the server looks up entities on the left-hand side of the join in its local storage and requests the entities on the right side of the join from other servers in the network.
+When a server receives a request from another server, it looks up the necessary entities to be joined locally and sends the reply back to the requesting server.
+Once the server processing a client request receives responses from the other servers on the network, it can send a final response back to the client.
+
+This results in the issue of stragglers we presented previously.
+Note that the common technique of issuing a request multiple times to avoid stragglers is not helpful in our scenario.
+As previously stated, our performance is bottlenecked by the available bandwidth.
+Stragglers exist in our system because of a lack of available bandwidth as opposed to variations in node performance (which is what multiple issuance is designed to mitigate).
+Furthermore, multiple issuance would place even stronger bandwidth pressure on our network, causing additional performance issues.
+Instead, we mitigate the effect of stragglers by adjusting the bandwidth allocation for each flow.
 

### 2014-08-04 paper/paper.tex
More related work

@@
 There has also been some work in exploiting software-defined networking for big data applications.
-Wang et al.~\cite{Wang2012} examined network traffic patterns in a Hadoop cluster and identified opportunities to combine network-aware job scheduling with dynamic routing configuration to improve job completion time.
-Xiong et al.~\cite{Xiong2014} examine the usefulness of software-defined networking in supporting distributed analytical queries.
+Wang et al.~\cite{Wang2012} examined network traffic patterns in a Hadoop~\cite{Shvachko2010} cluster and identified opportunities to combine network-aware job scheduling with dynamic routing configuration to improve job completion time.
+SmartJoin~\cite{Slagter2014} also examines network traffic patterns in Hadoop, specifically for join queries executed as MapReduce jobs.
+SmartJoin attempts to balance the overhead of sending tuples to remote servers to be processed with the performance gains of parallel query processing.
+These works hint at the benefits of network aware data processing systems, but focus on MapReduce jobs, while we we are interested in query processing over NoSQL database engines.
+
+CloudTPS~\cite{Wei2012} provides a join query processing layer on top of NoSQL stores which enforces ACID semantics for transactions.
+Our system does not currently consider transactional consistency, so this could be an interesting area to explore.
+R\"{o}diger et al.~\cite{Rodiger2014} examine the affects of data locality with respect to distributed query processing.
+Their Neo-Join algorithm is designed for efficient network usage by repartitioning data and scheduling network transfers to avoid cross-traffic.
+These techniques exert no control over the network, which may provide further opportunities for optimization.
+
+For example, Xiong et al.~\cite{Xiong2014} examine the usefulness of software-defined networking in supporting distributed analytical queries.
 Their workload consits of read-only SQL queries where query processing is bandwidth-intensive.

### 2014-08-05 src/CustomTopo.py
Adds the controller

@@ -6,10 +6,21 @@ from mininet.link import TCLink
 from mininet.cli import CLI
 import os
 import sys
+from mininet.net import Controller
 
 # adds the current dir i.e. src to system path to include sshd
 sys.path.append(os.path.dirname(__file__))
 from sshd import *
+
+class ControllerV1(Controller):
+    "Custom COntroller class to invoke our own forwarding.controller_v1"
+    def start(self):
+        "Starting controller v1"
+        self.pox = '%s/pox/pox.py' % os.environ['HOME']
+        self.cmd(self.pox, 'controller_v1 1> /mininet/src/out/pox.out 2>&1 &')
+    def stop(self):
+        "Stopping controller v1"
+        self.cmd('kill %' + self.pox)
 
 class DataCenter(Topo):
     "Simple Data Center Topology"
@@ -74,7 +85,7 @@ class TwoThreeTopo(Topo):
         sshdOpts = '-D -o UseDNS=no -u0'
         sshd
 
-default_linkopts = dict(bw=10, delay='5ms', loss=1, max_queue_size=1000, use_htb=True)
+default_linkopts = dict(bw=50, delay='5ms', loss=1, max_queue_size=1000, use_htb=True)
 topos = { '2s3h': ( lambda: TwoThreeTopo() ),
           'datacenter': (lambda: DataCenter(default_linkopts, default_linkopts, default_linkopts, 3)) }
 
@@ -95,10 +106,10 @@ def testDataCenter():
 
 def testSimpleDC():
     setLogLevel('info')
-    linkopts = dict(bw=10, delay='5ms', loss=1, max_queue_size=1000, use_htb=True)
+    linkopts = dict(bw=3, delay='5ms', use_htb=True)
     netopts = dict(host=CPULimitedHost, link=TCLink)
     topo = SimpleDC(linkopts, linkopts, 2);
-    net = Mininet(topo = topo,host=CPULimitedHost,link=TCLink)
+    net = Mininet(topo = topo,host=CPULimitedHost,link=TCLink, controller = ControllerV1)
     sshdOpts = '-D -o UseDNS=no -u0'
     sshd(net, switch=net['c1'], opts=sshdOpts)
 

### 2014-08-05 src/client.py

@@ -21,6 +21,8 @@ from joinHelpers import socket_read_n
 from joinHelpers import current_time_in_milli
 from joinHelpers import socket_send_data_by_json
 from joinHelpers import socket_recv_data_by_json
+from joinHelpers import LoadDifferentiator
+
 
 NUM_HOST = config['NUM_HOST']
 ID_LIST = []
@@ -31,14 +33,18 @@ BARRIER = BarrierSync()
 SEMAPHORE = threading.Semaphore()
 FINAL_JOIN_RESULT = []
 
+
 def loadItemIDs():
     with open(CSV_BASE + '/items.csv') as csvfile:
         reader = csv.reader(csvfile, delimiter=',', quotechar='\'',
                             doublequote=False, escapechar='\\')
         header = next(reader, None)
+        ld = LoadDifferentiator()
         for row in reader:
             id = row[0]
-            ID_LIST.append(int(id))
+            if(ld.include_huh(id)):
+                ID_LIST.append(int(id))
+    print('Total item to be read:' + str(len(ID_LIST)))
     # print ID_LIST
 
 def itemsJoin(host, port = 50007):
@@ -76,16 +82,17 @@ def joinItems():
 
 if __name__ == '__main__':
     loadItemIDs()
+    global_start_time = current_time_in_milli()
     try:
         joinItems()
     except:
         err(("Unexpected error:", sys.exc_info()[0]))
         raise
-    # sys.exit()
-    # barrierSynchronise(NUM_HOST + 1)
     BARRIER.sync(NUM_HOST + 1)
     time.sleep(1)
     total_len = 0
     for single_result in FINAL_JOIN_RESULT:
         total_len += len(single_result)
     err(('Operation completed, # of entries fetched:', total_len))
+    global_end_time = current_time_in_milli()
+    err(('Total time cost is: ', global_end_time - global_start_time,'ms'))

### 2014-08-05 src/controller_v1.py

@@ -1,0 +1,211 @@
+# Copyright 2011-2012 James McCauley
+#
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at:
+#
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""
+An L2 learning switch.
+
+It is derived from one written live for an SDN crash course.
+It is somwhat similar to NOX's pyswitch in that it installs
+exact-match rules for each flow.
+"""
+
+from pox.core import core
+import pox.openflow.libopenflow_01 as of
+from pox.lib.util import dpid_to_str
+from pox.lib.util import str_to_bool
+import time
+import sys
+
+
+sys.path.append('/mininet/src/')
+from joinHelpers import err
+
+log = core.getLogger()
+
+# We don't want to flood immediately when a switch connects.
+# Can be overriden on commandline.
+_flood_delay = 0
+
+class LearningSwitch (object):
+  """
+  The learning switch "brain" associated with a single OpenFlow switch.
+
+  When we see a packet, we'd like to output it on a port which will
+  eventually lead to the destination.  To accomplish this, we build a
+  table that maps addresses to ports.
+
+  We populate the table by observing traffic.  When we see a packet
+  from some source coming from some port, we know that source is out
+  that port.
+
+  When we want to forward traffic, we look up the desintation in our
+  table.  If we don't know the port, we simply send the message out
+  all ports except the one it came in on.  (In the presence of loops,
+  this is bad!).
+
+  In short, our algorithm looks like this:
+
+  For each packet from the switch:
+  1) Use source address and switch port to update address/port table
+  2) Is transparent = False and either Ethertype is LLDP or the packet's
+     destination address is a Bridge Filtered address?
+     Yes:
+        2a) Drop packet -- don't forward link-local traffic (LLDP, 802.1x)
+            DONE
+  3) Is destination multicast?
+     Yes:
+        3a) Flood the packet
+            DONE
[…]

### 2014-08-05 src/joinHelpers.py

@@ -1,3 +1,5 @@
+#!/usr/bin/env python
+
 import threading
 import threading
 import time
@@ -60,13 +62,14 @@ def socket_read_n(sock, n):
     return buf
 
 def err(listOfInput):
-    sys.stderr.write('DEBUG\n')
+    # sys.stderr.write('DEBUG\n')
     if type(listOfInput) is list or type(listOfInput) is tuple:
         for input in listOfInput:
             sys.stderr.write(str(input) + ' ')
     else:
         sys.stderr.write(str(listOfInput))
-    sys.stderr.write('\nDEBUG\n')
+    sys.stderr.write('\n')
+    # sys.stderr.write('\nDEBUG\n')
 
 class BarrierSync(object):
     def __init__(self):
@@ -83,3 +86,59 @@ class BarrierSync(object):
 
 def current_time_in_milli():
     return int(round(time.time() * 1000))
+
+
+# A quick & dirty implementation to have load differentiiation on servers
+
+class LoadDifferentiator(object):
+    def __init__(self):
+        # quick & dirty way
+        NUM_ITEMS_PER_HOST = 33722 / 4
+        self.num_host = int(config['NUM_HOST'])
+        self.buckets = dict()
+        self.max_buckets = dict()
+        for i in range(self.num_host):
+            self.buckets[i] = 0
+            self.max_buckets[i] = NUM_ITEMS_PER_HOST / (2**i)
+    # a duplicate of Store.bucket function
+    def storage_hash(self, key):
+        return hash(key) % self.num_host
+    def include_huh(self, item_id):
+        item_id_str = 'items:' + str(item_id)
+        server_id = self.storage_hash(item_id_str)
+        if(self.buckets[server_id] < self.max_buckets[server_id]):
+            self.buckets[server_id] += 1
+            return True
+        else:
+            return False
+
+    def assign_dscp(self, server_id):
+        """
+        4 return values           6, 16, 26, 36
+        correpsonding to queue    5, 3,  2,  1
+        """
+        if(server_id < self.num_host / 4.0):
+            return 6
+        if(server_id < self.num_host / 4.0 * 2):
+            return 16
+        if(server_id < self.num_host / 4.0 * 3 ):
+            return 26
+        else:
+            return 36
+    def dscp_to_queue_num(self, dscp):
+        return {
+            6: 5,
+            16: 3,
+            26: 2,
+            36: 1
+        }[dscp]
+
+
[…]

### 2014-08-05 src/scripts/queues_new.sh

@@ -1,0 +1,23 @@
+#!/bin/bash
+
+# Add queues with varying minimum bandwidth reservations to each switch port
+# queue 0 traffic will have priority over all other traffic
+
+for switch in $(sudo ovs-vsctl list-br); do
+    for port in $(sudo ovs-vsctl list-ports $switch); do
+        sudo ovs-vsctl -- set Port $port qos=@newqos -- \
+          --id=@newqos create QoS type=linux-htb other-config:max-rate=10000000 \
+          queues=0=@q0,1=@q1,2=@q2,3=@q3,4=@q4,5=@q5,6=@q6,7=@q7,8=@q8,9=@q9,10=@q10 -- \
+          --id=@q0 create Queue other-config:priority=0 -- \
+          --id=@q1 create Queue other-config:min-rate=1000000 other-config:priority=1 -- \
+          --id=@q2 create Queue other-config:min-rate=2000000 other-config:priority=1 -- \
+          --id=@q3 create Queue other-config:min-rate=3000000 other-config:priority=1 -- \
+          --id=@q4 create Queue other-config:min-rate=4000000 other-config:priority=1 -- \
+          --id=@q5 create Queue other-config:min-rate=5000000 other-config:priority=1 -- \
+          --id=@q6 create Queue other-config:min-rate=6000000 other-config:priority=1 -- \
+          --id=@q7 create Queue other-config:min-rate=7000000 other-config:priority=1 -- \
+          --id=@q8 create Queue other-config:min-rate=8000000 other-config:priority=1 -- \
+          --id=@q9 create Queue other-config:min-rate=9000000 other-config:priority=1 -- \
+          --id=@q10 create Queue other-config:min-rate=10000000 other-config:priority=1
+    done
+done

### 2014-08-05 src/scripts/start_topology.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+cp -f /mininet/src/controller_v1.py ~/pox/ext/.
+
+sudo python /mininet/src/CustomTopo.py

### 2014-08-05 src/scripts/test.sh

@@ -4,8 +4,11 @@ PROJECT_DIR='/mininet/src/'
 
 mkdir -p ${PROJECT_DIR}/out
 rm -rf  ~/.ssh/known_hosts
+
+bash /mininet/src/scripts/queues_new.sh
+
 for i in 10.0.0.{1..4}
 do
-    ssh ${i} "cd ${PROJECT_DIR}; python ./server.py >> out/server${i} 2>&1" &
+    ssh ${i} "cd ${PROJECT_DIR}; python ./server.py True >> out/server${i} 2>&1" &
     echo server started at ${i}
 done

### 2014-08-05 src/server.py

@@ -1,4 +1,5 @@
-# Echo server program
+#!/usr/bin/env python
+
 import socket
 import thread
 import sys
@@ -18,6 +19,7 @@ from joinHelpers import err
 from joinHelpers import socket_read_n
 from joinHelpers import socket_send_data_by_json
 from joinHelpers import socket_recv_data_by_json
+from joinHelpers import LoadDifferentiator
 from storage import *
 
 
@@ -27,7 +29,10 @@ HOST = ''    # Symbolic name meaning the local host
 # TODO: Read this port number form a configuration file
 PORT = int(config['SERVER_PORT']) # Arbitrary non-privileged port
 
-def getServerId():
+TRAFFIC_SCHEDULING = False
+LD = LoadDifferentiator()
+
+def get_server_id():
     # hostname = [(s.connect(('8.8.8.8', 80)), s.getsockname()[0], s.close()) for s in [socket.socket(socket.AF_INET, socket.SOCK_DGRAM)]][0][1]]
     # print(commands.getoutput("/sbin/ifconfig").split("\n")[1])
     ip = commands.getoutput("/sbin/ifconfig").split("\n")[1].split()[1][5:]
@@ -35,15 +40,18 @@ def getServerId():
     return rtn
 
 def fetch_users(sId, tuples, result_container, result_sema, barrier,
-                waitNum, partial_result):
+                waitNum, partial_result, origin_server_id):
     server_address = '10.0.0.' + str(sId + 1)
     # this socket is talking to another server to fetch relevant user data
-    sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
-    sock.connect((server_address, PORT))
+    s = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+    if TRAFFIC_SCHEDULING:
+        dscp = LD.assign_dscp(origin_server_id)
+        s.setsockopt(socket.SOL_IP, socket.IP_TOS, dscp << 2)
+    s.connect((server_address, PORT))
     allData = {config['OP'] : config['USERS_OP'],
                config['IDs'] : tuples}
-    socket_send_data_by_json(sock, allData)
-    partial_seller_result = socket_recv_data_by_json(sock)
+    socket_send_data_by_json(s, allData)
+    partial_seller_result = socket_recv_data_by_json(s)
     for key in partial_result:
         # err(partial_seller_result)
         # There is no replication at the moment so we can safely grab first item in [0]
[…]

### 2014-08-05 src/storage.py

@@ -158,10 +158,10 @@ class Query(object):
 if __name__ == '__main__':
     # Read a couple entities, run some queries with known IDs
     # and print the distribution of data between buckets
-    n = 4
+    n = config['NUM_HOST']
     # topo = StarTopo(n)
     store0 = Store(n, 3, 1)
     users = User(store0).read()
     items = Item(store0).read()
-    print(store0.get('items:533721')[1]['seller'])
+    # print(store0.get('items:533721')[1]['seller'])
     print store0.stats()

### 2014-08-05 paper/paper.tex
Expand introduction

@@
 This type of intelligent network behaviour is difficult to configure in a traditional network stack.
-However, with the advent of SDN, it is possible for a smart SDN controller to make global decisions based on the knowledge of higher-level application queries.
+With the advent of software-define networking (SDN), it is possible for a smart SDN controller to make global decisions based on the knowledge of higher-level application queries.
+Furthermore, the controller can dynamically adapt the network to improve query processing efficiency.
+
+As we discuss later, some existing query processing engines take into account network topology and link utilization when scheduling distributed queries.
+However, there is relatively little work expoloiting dynamic reconfiguration of the network to support query processing.
+A centralized network controller which is aware of the dynamics of query processing is able to make smart routing decisions to optimize performance.
+For example, flows for queries which are latency sensitive can be given priority over highly utilized links, while other flows can be routed to longer paths with lower utilization.
+This can be done dynamically based on the state of the network.
+Furthermore, the routing can be done using hardware switches as opposed to intelligent software routers, which minimizes latency and improves scalability.
+In this work, we examine one possible application of SDN for improving the throughput of distributed join queries.
 

### 2014-08-05 src/CustomTopo.py
Initial prototype of the system is completed

@@ -11,13 +11,18 @@ from mininet.net import Controller
 # adds the current dir i.e. src to system path to include sshd
 sys.path.append(os.path.dirname(__file__))
 from sshd import *
+from joinHelpers import config
+
 
 class ControllerV1(Controller):
-    "Custom COntroller class to invoke our own forwarding.controller_v1"
+    "Custom Controller class to invoke our own forwarding.controller_v1"
     def start(self):
         "Starting controller v1"
         self.pox = '%s/pox/pox.py' % os.environ['HOME']
-        self.cmd(self.pox, 'controller_v1 1> /mininet/src/out/pox.out 2>&1 &')
+        if(config['TRAFFIC_SCHEDULING'] == True):
+            self.cmd(self.pox, 'controller_v1 --traffic_scheduling=True 1> /mininet/src/out/pox.out 2>&1 &')
+        else:
+            self.cmd(self.pox, 'controller_v1 1> /mininet/src/out/pox.out 2>&1 &')
     def stop(self):
         "Stopping controller v1"
         self.cmd('kill %' + self.pox)
@@ -106,10 +111,10 @@ def testDataCenter():
 
 def testSimpleDC():
     setLogLevel('info')
-    linkopts = dict(bw=3, delay='5ms', use_htb=True)
-    netopts = dict(host=CPULimitedHost, link=TCLink)
+    linkopts = dict()
+    netopts = dict()
     topo = SimpleDC(linkopts, linkopts, 2);
-    net = Mininet(topo = topo,host=CPULimitedHost,link=TCLink, controller = ControllerV1)
+    net = Mininet(topo = topo, controller = ControllerV1)
     sshdOpts = '-D -o UseDNS=no -u0'
     sshd(net, switch=net['c1'], opts=sshdOpts)
 

### 2014-08-05 src/controller_v1.py

@@ -30,6 +30,7 @@ import sys
 
 sys.path.append('/mininet/src/')
 from joinHelpers import err
+from joinHelpers import LoadDifferentiator
 
 log = core.getLogger()
 
@@ -78,11 +79,16 @@ class LearningSwitch (object):
      flow goes out the appopriate port
      6a) Send the packet out appropriate port
   """
-  def __init__ (self, connection, transparent):
+  def __init__ (self, connection, transparent, traffic_scheduling = False):
     # sys.err.write(' XCLKJDSLFKJDSLKJFLDKSJ')
     # Switch we'll be adding L2 learning switch capabilities to
     self.connection = connection
     self.transparent = transparent
+
+    self.ld = LoadDifferentiator()
+    if(traffic_scheduling == 'True'):
+      traffic_scheduling = True
+    self.traffic_scheduling = traffic_scheduling
 
     # Our table
     self.macToPort = {}
@@ -174,11 +180,15 @@ class LearningSwitch (object):
         msg.match = of.ofp_match.from_packet(packet, event.port)
         msg.idle_timeout = 10
         msg.hard_timeout = 30
-        if(packet.type == 0x0800):
-          err((packet.payload.tos, packet.payload))
-          msg.actions.append(of.ofp_action_enqueue(port = port, queue_id = 10))
+        if self.traffic_scheduling:
+          if(packet.type == 0x0800):
+            queue_num = self.ld.dscp_to_queue_num(packet.payload.tos >> 2)
+            err((packet.payload.tos, queue_num, packet.payload))
+            msg.actions.append(of.ofp_action_enqueue(port = port, queue_id = queue_num))
+          else:
+            msg.actions.append(of.ofp_action_enqueue(port = port, queue_id = 6))
         else:
-          msg.actions.append(of.ofp_action_enqueue(port = port, queue_id = 10))
+          msg.actions.append(of.ofp_action_enqueue(port = port, queue_id = 0))
         msg.data = event.ofp # 6a
         self.connection.send(msg)
 
@@ -187,17 +197,18 @@ class l2_learning (object):
   """
   Waits for OpenFlow switches to connect and makes them learning switches.
   """
-  def __init__ (self, transparent):
+  def __init__ (self, transparent, traffic_scheduling):
     log.info('in init')
     core.openflow.addListeners(self)
     self.transparent = transparent
+    self.traffic_scheduling = traffic_scheduling
 
[…]

### 2014-08-05 src/joinHelpers.py

@@ -8,6 +8,8 @@ import json
 import struct
 
 config = {
+    'TRAFFIC_SCHEDULING' : False,
+
     'SERVER_PORT' : 50007,
     'NUM_HOST': 4,  # Update when when change topology
     'CSV_BASE': 'csv/',  # source of RUBiS data
@@ -126,12 +128,16 @@ class LoadDifferentiator(object):
         else:
             return 36
     def dscp_to_queue_num(self, dscp):
+        """
+        Converts dscp value to queue num
+        Default return value is queue # 6 which should be treated with highest priority
+        """
         return {
             6: 5,
             16: 3,
             26: 2,
             36: 1
-        }[dscp]
+        }.get(dscp, 6)
 
 
 # End of the quick & dirty implementation to have load differentiiation on servers
@@ -141,4 +147,4 @@ if __name__ == '__main__':
     ld = LoadDifferentiator()
     print ld.max_buckets
     print ld.assign_dscp(3)
-    print ld.dscp_to_queue_num(26)
+    print ld.dscp_to_queue_num(46)

### 2014-08-05 src/scripts/queues_new.sh

@@ -6,18 +6,18 @@
 for switch in $(sudo ovs-vsctl list-br); do
     for port in $(sudo ovs-vsctl list-ports $switch); do
         sudo ovs-vsctl -- set Port $port qos=@newqos -- \
-          --id=@newqos create QoS type=linux-htb other-config:max-rate=10000000 \
+          --id=@newqos create QoS type=linux-htb other-config:max-rate=1000000 \
           queues=0=@q0,1=@q1,2=@q2,3=@q3,4=@q4,5=@q5,6=@q6,7=@q7,8=@q8,9=@q9,10=@q10 -- \
           --id=@q0 create Queue other-config:priority=0 -- \
-          --id=@q1 create Queue other-config:min-rate=1000000 other-config:priority=1 -- \
-          --id=@q2 create Queue other-config:min-rate=2000000 other-config:priority=1 -- \
-          --id=@q3 create Queue other-config:min-rate=3000000 other-config:priority=1 -- \
-          --id=@q4 create Queue other-config:min-rate=4000000 other-config:priority=1 -- \
-          --id=@q5 create Queue other-config:min-rate=5000000 other-config:priority=1 -- \
-          --id=@q6 create Queue other-config:min-rate=6000000 other-config:priority=1 -- \
-          --id=@q7 create Queue other-config:min-rate=7000000 other-config:priority=1 -- \
-          --id=@q8 create Queue other-config:min-rate=8000000 other-config:priority=1 -- \
-          --id=@q9 create Queue other-config:min-rate=9000000 other-config:priority=1 -- \
-          --id=@q10 create Queue other-config:min-rate=10000000 other-config:priority=1
+          --id=@q1 create Queue other-config:min-rate=100000 other-config:priority=1 -- \
+          --id=@q2 create Queue other-config:min-rate=200000 other-config:priority=1 -- \
+          --id=@q3 create Queue other-config:min-rate=300000 other-config:priority=1 -- \
+          --id=@q4 create Queue other-config:min-rate=400000 other-config:priority=1 -- \
+          --id=@q5 create Queue other-config:min-rate=500000 other-config:priority=1 -- \
+          --id=@q6 create Queue other-config:min-rate=600000 other-config:priority=1 -- \
+          --id=@q7 create Queue other-config:min-rate=700000 other-config:priority=1 -- \
+          --id=@q8 create Queue other-config:min-rate=800000 other-config:priority=1 -- \
[…]

### 2014-08-05 src/scripts/test.sh

@@ -9,6 +9,6 @@ bash /mininet/src/scripts/queues_new.sh
 
 for i in 10.0.0.{1..4}
 do
-    ssh ${i} "cd ${PROJECT_DIR}; python ./server.py True >> out/server${i} 2>&1" &
+    ssh ${i} "cd ${PROJECT_DIR}; python ./server.py >> out/server${i} 2>&1" &
     echo server started at ${i}
 done

### 2014-08-05 src/server.py

@@ -10,6 +10,7 @@ import os
 import traceback
 import ast
 import struct
+import re
 
 # adds the current dir i.e. src to system path
 sys.path.append(os.path.dirname(__file__))
@@ -31,6 +32,8 @@ PORT = int(config['SERVER_PORT']) # Arbitrary non-privileged port
 
 TRAFFIC_SCHEDULING = False
 LD = LoadDifferentiator()
+
+SERVER_IP_REGEX = re.compile('.*?\d+\.\d+\.\d+\.(\d+).*')
 
 def get_server_id():
     # hostname = [(s.connect(('8.8.8.8', 80)), s.getsockname()[0], s.close()) for s in [socket.socket(socket.AF_INET, socket.SOCK_DGRAM)]][0][1]]
@@ -63,10 +66,18 @@ def fetch_users(sId, tuples, result_container, result_sema, barrier,
     s.close()
     barrier.sync(waitNum)
 
-def serve_user_op(client_socket, storage, client_ids):
+def serve_user_op(client_socket, client_addr, storage, client_ids):
     rtn = dict()
     for client_id in client_ids:
         rtn[client_id] = storage.get(client_id)
+    if TRAFFIC_SCHEDULING:
+        matched = SERVER_IP_REGEX.match(str(client_addr))
+        assert matched != None
+        # The 'client' is actually a server with an ID in this case
+        client_server_id = int(matched.group(1)) - 1
+        # err((client_server_id, client_addr))
+        dscp = LD.assign_dscp(client_server_id)
+        client_socket.setsockopt(socket.SOL_IP, socket.IP_TOS, dscp << 2)
     socket_send_data_by_json(client_socket, rtn)
 
 def handleJoinItemUserOp(clientsocket, addr, storage, serverId, item_ids):
@@ -116,7 +127,7 @@ def handleJoinItemUserOp(clientsocket, addr, storage, serverId, item_ids):
     return result
 
 
-def serve(client_socket, addr, storage, serverId):
+def serve(client_socket, addr, storage, server_id):
     """ Each serve call is a new thread that serves requests """
     lenBuf = client_socket.recv(4)
     dataLen = struct.unpack('>L', lenBuf)[0]
@@ -126,11 +137,14 @@ def serve(client_socket, addr, storage, serverId):
     ids = dataDict[config['IDs']]
     op = dataDict[config['OP']]
     if(op == config['JOIN_ITEM_USER_OP']):
[…]

### 2014-08-06 src/scripts/start_topology.sh
Ensure pox ext directory is created

@@ -1,5 +1,6 @@
 #!/bin/bash
 
+mkdir -p ~/pox/ext
 cp -f /mininet/src/controller_v1.py ~/pox/ext/.
 
 sudo python /mininet/src/CustomTopo.py

### 2014-08-06 src/CustomTopo.py
Fix topology to run standalone

@@ -6,10 +6,11 @@ from mininet.link import TCLink
 from mininet.cli import CLI
 import os
 import sys
+import inspect
 from mininet.net import Controller
 
 # adds the current dir i.e. src to system path to include sshd
-sys.path.append(os.path.dirname(__file__))
+sys.path.append(os.path.dirname(os.path.abspath(inspect.getsourcefile(lambda _: None))))
 from sshd import *
 from joinHelpers import config
 
@@ -91,8 +92,11 @@ class TwoThreeTopo(Topo):
         sshd
 
 default_linkopts = dict(bw=50, delay='5ms', loss=1, max_queue_size=1000, use_htb=True)
-topos = { '2s3h': ( lambda: TwoThreeTopo() ),
-          'datacenter': (lambda: DataCenter(default_linkopts, default_linkopts, default_linkopts, 3)) }
+topos = {
+  '2s3h': lambda: TwoThreeTopo(),
+  'datacenter': lambda: DataCenter(default_linkopts, default_linkopts, default_linkopts, 3),
+  'simple': lambda: SimpleDC({}, {}, 2),
+}
 
 def testDataCenter():
     setLogLevel('info')

### 2014-08-06 src/controller_v1.py
removes the timeout for the openflow rules & make the controllers use packet.payload.tos as the queue number

@@ -30,7 +30,6 @@ import sys
 
 sys.path.append('/mininet/src/')
 from joinHelpers import err
-from joinHelpers import LoadDifferentiator
 
 log = core.getLogger()
 
@@ -85,7 +84,6 @@ class LearningSwitch (object):
     self.connection = connection
     self.transparent = transparent
 
-    self.ld = LoadDifferentiator()
     if(traffic_scheduling == 'True'):
       traffic_scheduling = True
     self.traffic_scheduling = traffic_scheduling
@@ -178,15 +176,16 @@ class LearningSwitch (object):
                   (packet.src, event.port, packet.dst, port))
         msg = of.ofp_flow_mod()
         msg.match = of.ofp_match.from_packet(packet, event.port)
-        msg.idle_timeout = 10
-        msg.hard_timeout = 30
+        # msg.idle_timeout = 10
+        # msg.hard_timeout = 30
         if self.traffic_scheduling:
           if(packet.type == 0x0800):
-            queue_num = self.ld.dscp_to_queue_num(packet.payload.tos >> 2)
+            # queue_num = self.ld.dscp_to_queue_num(packet.payload.tos >> 2)
+            queue_num = (packet.payload.tos >> 2)
             err((packet.payload.tos, queue_num, packet.payload))
             msg.actions.append(of.ofp_action_enqueue(port = port, queue_id = queue_num))
           else:
-            msg.actions.append(of.ofp_action_enqueue(port = port, queue_id = 6))
+            msg.actions.append(of.ofp_action_enqueue(port = port, queue_id = 0))
         else:
           msg.actions.append(of.ofp_action_enqueue(port = port, queue_id = 0))
         msg.data = event.ofp # 6a

### 2014-08-06 src/CustomTopo.py
Modified CustomTopo so now the client runs within mininet and it is connected to the core switch

@@ -69,6 +69,9 @@ class SimpleDC(Topo):
             for k in range(fanout):
                 host = self.addHost('h' + str(fanout * i + (k + 1)))
                 self.addLink(aggregationSwitch, host, **linkopts2)
+        host = self.addHost('z_client' + str(1))
+        self.addLink(coreSwitch, host)
+
 
 class TwoThreeTopo(Topo):
     """

### 2014-08-06 src/joinHelpers.py

@@ -8,7 +8,7 @@ import json
 import struct
 
 config = {
-    'TRAFFIC_SCHEDULING' : False,
+    'TRAFFIC_SCHEDULING' : True,
 
     'SERVER_PORT' : 50007,
     'NUM_HOST': 4,  # Update when when change topology
@@ -116,29 +116,16 @@ class LoadDifferentiator(object):
 
     def assign_dscp(self, server_id):
         """
-        4 return values           6, 16, 26, 36
         correpsonding to queue    5, 3,  2,  1
         """
         if(server_id < self.num_host / 4.0):
-            return 6
+            return 8
         if(server_id < self.num_host / 4.0 * 2):
-            return 16
+            return 3
         if(server_id < self.num_host / 4.0 * 3 ):
-            return 26
+            return 2
         else:
-            return 36
-    def dscp_to_queue_num(self, dscp):
-        """
-        Converts dscp value to queue num
-        Default return value is queue # 6 which should be treated with highest priority
-        """
-        return {
-            6: 5,
-            16: 3,
-            26: 2,
-            36: 1
-        }.get(dscp, 6)
-
+            return 1
 
 # End of the quick & dirty implementation to have load differentiiation on servers
 

### 2014-08-06 src/scripts/run_client.sh
adds new script to invoke client in scripts, adds partial evaluation results

@@ -1,0 +1,28 @@
+#!/bin/bash
+
+if [ $# -ne 3 ]
+then
+    echo "Usage: ${0} client_ip bandwidth experiment_name"
+    echo "Example Usage:"
+    echo "    ${0} 10.0.0.5 1mbps norm"
+    exit 66
+fi
+
+PROJECT_DIR='/mininet/src/'
+result_dir="${PROJECT_DIR}/result/${2}/${3}/"
+mkdir -p ${result_dir}
+rm -rf  ~/.ssh/known_hosts
+
+for i in {0..5};
+do
+    ssh ${1} "cd ${PROJECT_DIR}; ./client.py  > ${result_dir}/client_${i}.out 2>&1"
+    sleep 5
+done
+
+
+# Make some sound so we know this is finished!
+echo -ne '\007'
+echo -ne '\007'
+echo -ne '\007'
+echo -ne '\007'
+echo -ne '\007'

### 2014-08-07 paper/paper.tex
Add additional introduction paragraph

@@
 Furthermore, the routing can be done using hardware switches as opposed to intelligent software routers, which minimizes latency and improves scalability.
+
 In this work, we examine one possible application of SDN for improving the throughput of distributed join queries.
+Specifically, we aim to schedule network traffic between servers processing distributed joins to optimize query completion time.
+As identified above, improved scheduling can result in increased query performance.
+We contribute a simple but effective method of flow scheduling in this use case.
+We also identify several ways our work can be complemented with other techniques and extended to support alternative query models.
 

### 2014-08-07 plot_src/confidence_interval_calculator.py
Add more results & two initial plots

@@ -1,0 +1,34 @@
+#!/usr/bin/python -B
+
+import sys
+import math
+
+# for MicroFuge the default T_TEST Value is set to the following
+# df ==4 & 95% confidence interval
+T_TEST = 2.776
+
+# for now, suppose there is no missing value
+# TODO Add missing value support like my java implementation
+# return an array of size two
+# the first value is the mean, the second value is the stderr
+def get_conf_int(array, debug = False):
+    rtn = []
+    sum = 0.0
+    for i in range(len(array)):
+        sum += float(array[i])
+    std_mean = sum / len(array)
+    rtn.append(std_mean)
+    std_derivation = 0.0
+    for i in range(len(array)):
+        std_derivation += ((std_mean - float(array[i])) * (std_mean - float(array[i])))
+    std_derivation /= (len(array) - 1)
+    std_derivation = math.sqrt(std_derivation)
+    std_derivation = T_TEST * (std_derivation / math.sqrt(len(array)))
+    rtn.append(std_derivation)
+    return rtn
+
+# for now, mannually verified with confident interval I wrote in java
+def xcui_test():
+    a=[45,55,67,45,68,79,98,87,84,82]
+    rtn = get_conf_int(a)
+    print rtn

### 2014-08-07 plot_src/save_fig.py

@@ -1,0 +1,62 @@
+#!/usr/bin/python -B
+
+# Thank to Jessica B. Hamrick
+# I copied this code from her github which is linked from her website at:
+# http://www.jesshamrick.com/2012/09/03/saving-figures-from-pyplot/
+import os
+import sys
+
+def save(user_plt, path, ext='png', close=True, verbose=True):
+    """Save a figure from pyplot.
+
+    Parameters
+    ----------
+    user_plt: matplotlib.pyplot
+        The pyplot object so we can use it to draw our graph
+
+    path : string
+        The path (and filename, without the extension) to save the
+        figure to.
+
+    ext : string (default='png')
+        The file extension. This must be supported by the active
+        matplotlib backend (see matplotlib.backends module).  Most
+        backends support 'png', 'pdf', 'ps', 'eps', and 'svg'.
+
+    close : boolean (default=True)
+        Whether to close the figure after saving.  If you want to save
+        the figure multiple times (e.g., to multiple formats), you
+        should NOT close it in between saves or you will have to
+        re-plot it.
+
+    verbose : boolean (default=True)
+        Whether to print information about when and where the image
+        has been saved.
+
+    """
+
+    # Extract the directory and filename from the given path
+    directory = os.path.split(path)[0]
+    filename = "%s.%s" % (os.path.split(path)[1], ext)
+    if directory == '':
+        directory = '.'
+
+    # If the directory does not exist, create it
+    if not os.path.exists(directory):
+        os.makedirs(directory)
+
+    # The final path to save to
+    savepath = os.path.join(directory, filename)
+
+    if verbose:
+        sys.stderr.write("Saving figure to '%s'..." % savepath)
+
+    # Actually save the figure
+    user_plt.savefig(savepath)
+
+    # Close it
+    if close:
+        user_plt.close()
+
+    if verbose:
+        sys.stderr.write("Done\n")

### 2014-08-07 plot_src/simple_dc_closer_8.py

@@ -1,0 +1,90 @@
+#!/usr/bin/env python
+# a stacked bar plot with errorbars
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+import re
+from os.path import expanduser
+import confidence_interval_calculator as CI
+import save_fig as saver
+
+
+def get_overall_arr(key_word, data_dir):
+    rtn = []
+    for i in range(1,6):
+        data = open(data_dir + 'client_' + (str(i)) + '.out' , "r")
+        for line in data:
+            if(re.search(key_word, line)):
+                tokens = line.split(" ")
+                rtn.append(float(tokens[len(tokens) - 2]))
+    return rtn
+
+PATTERN = [ "/" , "\\" ,"o", "*", "O", "x", ".", '-', "|" , "+"]
+base_dir = '/mininet/src/result/1mbps/'
+data_dirs = [base_dir + 'norm/',
+             base_dir + 'scheduling_queue_8_for_h1/']
+
+
+
+base_mean = []
+base_err = []
+
+# return values will be filled into rtn_mean, rtn_std
+def get_results(key_word):
+    for i in range(len(data_dirs)):
+        overall_arr = get_overall_arr(key_word, data_dirs[i])
+        stdmean_stderr = CI.get_conf_int(overall_arr)
+        base_mean.append(int(stdmean_stderr[0] / 100) / 10.0)
+        base_err.append(int(stdmean_stderr[1] / 100) / 10.0)
+
+N = 4
+ind = np.arange(N)    # the x locations for the groups
+width = 0.22      # the width of the bars
+
+### plotting now
+# this clears the current figure
+
+plt.clf()
+plt.figure(None, figsize=(8,5), dpi = 300)
+
+ind = ind + 0.3
+
+final_base = []
+final_base_err = []
+final_comp = []
+final_comp_err = []
+
+for i in reversed(range(4)):
+    base_mean = []
+    base_err = []
+    get_results('10.0.0.' + str(i+1))
+    final_base.append(base_mean[0])
+    final_base_err.append(base_err[0])
+    final_comp.append(base_mean[1])
+    final_comp_err.append(base_err[1])
+
+
+p0 = plt.bar(ind + width * 0,  final_base, width, color='w',
+             yerr=final_base_err, ecolor='b',
+             edgecolor='black', hatch=PATTERN[0])
+
[…]

### 2014-08-07 plot_src/simple_dc_overall.py

@@ -1,0 +1,93 @@
+#!/usr/bin/env python
+# a stacked bar plot with errorbars
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+import re
+from os.path import expanduser
+import confidence_interval_calculator as CI
+import save_fig as saver
+
+
+def get_overall_arr(key_word, data_dir):
+    rtn = []
+    for i in range(1,6):
+        data = open(data_dir + 'client_' + (str(i)) + '.out' , "r")
+        for line in data:
+            if(re.search(key_word, line)):
+                tokens = line.split(" ")
+                rtn.append(float(tokens[len(tokens) - 3]))
+    return rtn
+
+PATTERN = [ "/" , "\\" ,"o", "*", "O", "x", ".", '-', "|" , "+"]
+base_dir = '/mininet/src/result/1mbps/'
+data_dirs = [base_dir + 'norm/',
+             base_dir + 'scheduling_queue_5_for_h1/',
+             base_dir + 'scheduling_queue_6_for_h1/',
+             base_dir + 'scheduling_queue_7_for_h1/',
+             base_dir + 'scheduling_queue_8_for_h1/',
+             base_dir + 'scheduling_queue_9_for_h1/',
+             base_dir + 'scheduling_queue_10_for_h1/']
+
+
+rtn_mean = []
+rtn_std = []
+
+# return values will be filled into rtn_mean, rtn_std
+def get_results():
+    for i in range(len(data_dirs)):
+        overall_arr = get_overall_arr('Total time', data_dirs[i])
+        stdmean_stderr = CI.get_conf_int(overall_arr)
+        rtn_mean.append(int(stdmean_stderr[0] / 100) / 10.0)
+        rtn_std.append(int(stdmean_stderr[1] / 100) / 10.0)
+
+
+
+get_results()
+print rtn_mean
+
+base_mean = [rtn_mean[0], rtn_mean[0], rtn_mean[0], rtn_mean[0], rtn_mean[0], rtn_mean[0]]
+base_err = [rtn_std[0], rtn_std[0], rtn_std[0], rtn_std[0], rtn_std[0], rtn_std[0]]
+compare_mean = rtn_mean[1:]
+compare_err = rtn_std[1:]
+
+N = 6
+ind = np.arange(N)    # the x locations for the groups
+width = 0.22      # the width of the bars
+
+### plotting now
+# this clears the current figure
+
+plt.clf()
[…]

### 2014-08-07 paper/paper.tex
Adds initial version of the evaluation

@@
 As we discuss later, some existing query processing engines take into account network topology and link utilization when scheduling distributed queries.
-However, there is relatively little work expoloiting dynamic reconfiguration of the network to support query processing.
+However, there is relatively little work exploiting dynamic reconfiguration of the network to support query processing.
 A centralized network controller which is aware of the dynamics of query processing is able to make smart routing decisions to optimize performance.
@@
 Those nodes will scan their local list of entities and apply the necessary filters (\texttt{price > 20}).
-Each node then forwards the necessary foreign keys to nodes holding entites of the class on the right side of the join (\texttt{User}).
+Each node then forwards the necessary foreign keys to nodes holding entities of the class on the right side of the join (\texttt{User}).
 These nodes are then able to look up entities on the other side of the join and results can be passed back to the client.
@@
 
-Simple approaches to bandwith allocation such as max-min fairness pose significant problems in our settings.
+Simple approaches to bandwidth allocation such as max-min fairness pose significant problems in our settings.
 If we were to use fair share allocation, then the node with the largest share of the query results, bottlenecked by available bandwidth, would become a straggler.
@@
 R\"{o}diger et al.~\cite{Rodiger2014} examine the affects of data locality with respect to distributed query processing.
-Their Neo-Join algorithm is designed for efficient network usage by repartitioning data and scheduling network transfers to avoid cross-traffic.
+Their Neo-Join algorithm is designed for efficient network usage by re-partitioning data and scheduling network transfers to avoid cross-traffic.
 These techniques exert no control over the network, which may provide further opportunities for optimization.
@@
 For example, Xiong et al.~\cite{Xiong2014} examine the usefulness of software-defined networking in supporting distributed analytical queries.
-Their workload consits of read-only SQL queries where query processing is bandwidth-intensive.
+Their workload consists of read-only SQL queries where query processing is bandwidth-intensive.
 They construct a global query optimizer which decides on join order and how query results should be passed between sites.
@@
 
+Each experiment is performed by 5 independent runs and the results are used to
+compute the data in the following sections. The 95th percentile confidence
+interval is shown whenever appropriate.
+\subsection{Micro Benchmark}
+\begin{figure}
+    \centering
+    \includegraphics[scale=2.5]{figures/Join_micro_benchmark.png}
+    \caption{Micro benchmark setup}\label{fig:micro_benchmark}
+\end{figure}
+First, we want to demonstrate the proof-of-concept of our design by evaluating
+our system with the simple topology shown in figure
+\ref{fig:micro_benchmark}. We engineered the client query such that each server
+holds a different portion of the relevant data for the join operation,
+i.e. server 1 holds a quarter of the relevant \texttt{item} data and server 2,
+3, 4 holds 1/16, 1/32, 1/64 of the relevant \texttt{item} data
+respectively. The total relevant data sizes of \texttt{item} and \texttt{user}
+are about 9 MB and 2 MB respectively. In order to satisfy our assumption where
+the network is the bottleneck, we set the capacity of each link in the topology
+to 1Mbps.  The client is connected directly to the core switch with the same
+link capacity.
+
+In figure \ref{fig:micro_benchmark_total_time}, the red bars represent our
+system's baseline performance where the network traffic scheduling is turned
+off, i.e. without SDN. The average completion time without scheduling for each
+client query is over 180 seconds with a high variation. This kind of
+unpredictable performance is undesired in any datacenters. To provide
+predictable performance with our scheduling mechanism, we must first understand
+the characteristics of the costs for each sub query of the join operation. In
+our case it is easy to predict the costs because we assigned different portions
+of relevant data to each server, however, it is also easy in real deployment
+since the statistics of the server is usually available. Recall that by our
+design, server 1 is responsible to transmit the largest flows from and to other
+servers and the client.  Therefore, our scheduling must take this information
+into consideration and ensure that server 1 have the largest minimum bandwidth
+guarantee for this particular operation. Though we want server 1 to transfer
+its data as fast as possible, we do not want to starve other
+servers. Therefore, we must also have reasonable minimum bandwidth guarantees
+for other servers. Through experimentation, we find the most suitable minimum
+bandwidth guarantees on the 1 Mbps link for servers 2, 3, and 4 are 0.3 Mbps,
+0.2 Mbps and 0.1 Mbps respectively. We fixed these parameters and varied the
+minimum bandwidth guarantee for the server 1 to study the network
+behaviours. As shown by the yellow bars in figure
+\ref{fig:micro_benchmark_total_time}, there is noticeable performance gain when
+we increase the minimum bandwidth guarantee of server 1 from 0.5 Mbps to 0.8
+Mbps. It shows that it is important to favour the elephant flows during a join
+operation to minimize the overall completion time. We also want to highlight
+that we see diminished return as we increase the guarantee to 0.9 Mbps or even
+1.0 Mbps because this may significantly hurt other flows' completion time which
+result in slower execution overall. In the best case, with 0.8 minimum
+bandwidth guarantee for server 1, the average completion time of the client
+query is 154 seconds which gives us almost 15\% performance gain over the
+baseline implementation.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=0.47]{figures/simple_topo_overall.png}
+    \caption{Query completion time}\label{fig:micro_benchmark_total_time}
+\end{figure}
+
+Next, we want to take a closer look at the performance results to understand
+why network scheduling helps. Figure \ref{fig:sub_query_completion_8} shows the
+sub query completion time for each server. The per server completion is
+calculated by the client after it fetched all the required data from each
+server. The white bars on the left is generated when the traffic scheduling is
+turned off. We see that server 1 takes the most time to complete and server 4
+takes the least time. This is expected due to the relevant data
+distribution. We can also see that if we turned on our network traffic
+scheduling mechanisms, the completion time for server 3 and server 4 slowed
+down quite a bit. This is because most of the bandwidth capacity will be
+allocated to help server 1 to finish its task as soon as possible to minimize
+final query completion time. The end result is of traffic scheduling is to
+average out the sub query completion time which in turn reduces the overall
+query execution time. We see that our experimental results match with our
+expectations.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=0.47]{figures/simple_topo_closer_8.png}
+    \caption{Sub query completion time}\label{fig:sub_query_completion_8}
+\end{figure}
+
 \section{Future Work}
@@
 There are many opportunities to combine this work with appropriate key placement strategies for further performance optimizations.
-For heterogenous networks, a key placement strategy which is network and workload aware may provide significant performance advantages.
+For heterogeneous networks, a key placement strategy which is network and workload aware may provide significant performance advantages.
 Our current query execution model is very simplistic and could be expanded to support richer queries and distributed index data structures.

### 2014-08-07 src/scripts/queues_new.sh
Modified scripts so queue_new.sh can take bandwidth capacity as input

@@ -1,23 +1,31 @@
 #!/bin/bash
 
+
+if [ ${#} -ne 1 ]
+then
+    echo "Usage: ${0} bandwidth_in_bps"
+fi
+
+# total_bandwidth=1000000
+total_bandwidth=${1}
+
 # Add queues with varying minimum bandwidth reservations to each switch port
 # queue 0 traffic will have priority over all other traffic
-
 for switch in $(sudo ovs-vsctl list-br); do
     for port in $(sudo ovs-vsctl list-ports $switch); do
         sudo ovs-vsctl -- set Port $port qos=@newqos -- \
-          --id=@newqos create QoS type=linux-htb other-config:max-rate=1000000 \
+          --id=@newqos create QoS type=linux-htb other-config:max-rate=${total_bandwidth} \
           queues=0=@q0,1=@q1,2=@q2,3=@q3,4=@q4,5=@q5,6=@q6,7=@q7,8=@q8,9=@q9,10=@q10 -- \
           --id=@q0 create Queue other-config:priority=0 -- \
-          --id=@q1 create Queue other-config:min-rate=100000 other-config:priority=1 -- \
-          --id=@q2 create Queue other-config:min-rate=200000 other-config:priority=1 -- \
-          --id=@q3 create Queue other-config:min-rate=300000 other-config:priority=1 -- \
-          --id=@q4 create Queue other-config:min-rate=400000 other-config:priority=1 -- \
-          --id=@q5 create Queue other-config:min-rate=500000 other-config:priority=1 -- \
-          --id=@q6 create Queue other-config:min-rate=600000 other-config:priority=1 -- \
-          --id=@q7 create Queue other-config:min-rate=700000 other-config:priority=1 -- \
-          --id=@q8 create Queue other-config:min-rate=800000 other-config:priority=1 -- \
-          --id=@q9 create Queue other-config:min-rate=900000 other-config:priority=1 -- \
-          --id=@q10 create Queue other-config:min-rate=1000000 other-config:priority=1
+          --id=@q1 create Queue other-config:min-rate=$((total_bandwidth/10*1)) other-config:priority=1 -- \
+          --id=@q2 create Queue other-config:min-rate=$((total_bandwidth/10*2)) other-config:priority=1 -- \
+          --id=@q3 create Queue other-config:min-rate=$((total_bandwidth/10*3)) other-config:priority=1 -- \
[…]

### 2014-08-07 src/scripts/test.sh

@@ -1,11 +1,18 @@
 #!/bin/bash
 
+if [ ${#} -ne 1 ]
+then
+    echo "Usage: ${0} link_bandwidth_in_bsp"
+fi
+
+total_bandwidth=${1}
+
 PROJECT_DIR='/mininet/src/'
 
 mkdir -p ${PROJECT_DIR}/out
 rm -rf  ~/.ssh/known_hosts
 
-bash /mininet/src/scripts/queues_new.sh
+bash /mininet/src/scripts/queues_new.sh ${total_bandwidth}
 
 for i in 10.0.0.{1..4}
 do

### 2014-08-07 paper/paper.tex
Report edits

--- 2014-08-07 paper/paper.tex
@@
 
+Each experiment is performed 5 times with the results are used to compute the data in the following sections.
+The 95th percentile confidence interval is shown where appropriate.
+
+\subsection{Micro Benchmark}
+\begin{figure}
+    \centering
+    \includegraphics[scale=2.5]{figures/Join_micro_benchmark.png}
+    \caption{Micro benchmark setup}\label{fig:micro_benchmark}
+\end{figure}
+
+First, we want to demonstrate the proof-of-concept of our design by evaluating our system with the simple topology shown in Figure~\ref{fig:micro_benchmark}.
+We engineered the client query such that each server holds a different portion of the relevant data for the join operation.
+That is, server 1 holds $1/4$ of the relevant item data and server 2, 3, and 4 hold $1/16$, $1/32$, $1/64$ of the relevant item data respectively.
+The total sizes of the item and user data are approximately 9 MB and 2 MB\@.
+In order to satisfy our assumption where the network is the bottleneck, we set the capacity of each link in the topology to 1 Mbps.
+The client is connected directly to the core switch with a link of the same capacity.
+
+In Figure~\ref{fig:micro_benchmark_total_time}, the red bars represent our system's baseline performance without our scheduling mechanism is turned off.
+The average completion time without scheduling for each client query is over 180 seconds and also shows significant variation.
+To provide predictable performance with our scheduling mechanism, we must first understand the characteristics of the costs for each component of the join operation.
+In our case it is easy to predict the costs because we assigned different portions of relevant data to each server.
+However, it is also easy in a real deployment scenario as the data distribution among servers is typically available or can be easily computed.
+Recall that by design, server 1 is responsible for transmitting the largest flows between other servers and the client. 
+Therefore, our scheduling must take this information into consideration and ensure that server 1 has the largest minimum bandwidth guarantee for this particular operation.
+
+Though we want server 1 to transfer its data as fast as possible, we do not want to starve other servers.
+Therefore, we must also have reasonable minimum bandwidth guarantees for other servers.
+Through experimentation, we found the most suitable minimum bandwidth guarantees on the 1 Mbps link for servers 2, 3, and 4 are 0.3 Mbps, 0.2 Mbps and 0.1 Mbps respectively.
+We fixed these parameters and varied the minimum bandwidth guarantee for server 1 to study the network behaviour.
+As shown by the right bars in Figure~\ref{fig:micro_benchmark_total_time}, there is a noticeable performance gain when we increase the minimum bandwidth guarantee of server 1 from 0.5 Mbps to 0.8 Mbps.
+This shows that it is important to favour larger flows during a join operation to minimize the overall completion time.
+We also want to highlight that we see diminishing returns as we further increase the guarantee.
+This is caused by a significant increase in the completion time for other flows, hurting the overall performance.
+In the best case, with 0.8 Mbps minimum bandwidth guarantee for server 1, the average completion time of the client query is 154 seconds which gives us almost 15\% performance gain over the baseline implementation.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=0.47]{figures/simple_topo_overall.png}
+    \caption{Query completion time}\label{fig:micro_benchmark_total_time}
+\end{figure}
+
+Next, we further examine the performance results to understand why network scheduling helps.
+Figure~\ref{fig:sub_query_completion_8} shows the subquery completion time for each server.
+The per-server completion is calculated by the client after it fetches all the required data from each server.
+The bars on the left are generated when traffic scheduling is disabled.
+We see that server 1 takes the most time to complete and server 4 takes the least time.
+This is expected due to the data distribution we defined above.
+We can also see that when we enabled our network traffic scheduling mechanisms, the completion time for server 3 and server 4 slowed down slightly.
+This is because most of the bandwidth will be allocated to help server 1 to finish its task as soon as possible.
+However, The end result is that traffic scheduling reduces the average subquery completion time, which in turn reduces the overall query execution time.
+We see that our experimental results match with our expectations.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=0.47]{figures/simple_topo_closer_8.png}
+    \caption{Sub query completion time}\label{fig:sub_query_completion_8}
+\end{figure}
+
 \section{Future Work}

### 2014-08-07 paper/paper.tex
Additional future work paragraph

@@
 
+While our results are promising, we would like to take a more principled approach to determine the theoretical gains possible for applying this technique in order to clearly identify potential bottlenecks.
+We are also interested in performing a more thorough evaluation on a larger scale with more realistic conditions.
+We feel that there is potential to significantly improve the performance of our current approach.
+For example, we currently buffer all results at each server before sending a reply.
+Streaming results back to the client as they are received may result in higher link utilization and increased efficiency.
+
 There are many opportunities to combine this work with appropriate key placement strategies for further performance optimizations.

### 2014-08-07 paper/paper.tex
Add first pass at conclusion

@@
 
-Recent trends have shown a proliferation of databases termed ``NoSQL'' which frequently eschew traditional database concepts such as referential integrity and ACID transactions for the sake of performance.
+Recent trends have shown a proliferation of databases termed ``NoSQL'' which frequently eschew traditional database concepts such as referential integrity, ACID transactions, and high-level query languages for the sake of performance.
 However, applications using these databases still have a need to perform complex queries.
@@
 
+In this work, we examined opportunties to apply software-defined network techniques to provide more efficient query processing over NoSQL databases.
+We demonstrated a simple technique using smart network scheduling to improve the performance of distributed joins.
+Our technique provides a significant performance improvement over uninformed switching techniques.
+
+We also presented several opportunties to further exploit of software-defined network techniques to improve query processing.
+As software-defined networking becomes increasingly common in datacentres, we feel similar techniques will become increasingly prevalent.
+Little work has been done to exploit software-defined networking in the context of NoSQL databases and distributed data management.
+
 \let\theOLDbibliography\thebibliography\renewcommand{\thebibliography}[1]{\theOLDbibliography{#1}%

### 2014-08-08 paper/paper.tex
Add related work section on partitioning

@@
 
+The traditional approach to optimizing network usage in distributed query processing is to apply smart key placement strategies to minimize the number of distributed transactions, and also the amount of data which must be transferred between nodes.
+For example, Schism~\cite{Curino2010} uses graph partitioning techniques to reduce the cost of distributed transactions by up to 30\%.
+Vila\c{c}a et al.~\cite{Vilaca2011} specifically target key-value stores and aim to improve locality in a multidimensional space of tags (such as foreign keys) applied to data items.
+We expect partitioning techniques such as these to be complementary to our work.
+
 There has also been some work in exploiting software-defined networking for big data applications.

### 2014-08-08 paper/paper.tex
Added more text to future work and conclusion

@@
 
+There are other advanced join operations which may benefit from software-defined
+networking. Ntarmos et al.~\cite{Ntarmos2014} uses novel indexing techniques to
+improve the performance of top-k joins for NoSQL databases which in turn
+reduces both network latency and network traffics. This work did not explicit
+use software-defined networking, however the amount of network traffic is one
+important metric in their evaluation which hints possible network related
+work in this domain.
+
 \section{Evaluation}
@@
 
-While our results are promising, we would like to take a more principled approach to determine the theoretical gains possible for applying this technique in order to clearly identify potential bottlenecks.
-We are also interested in performing a more thorough evaluation on a larger scale with more realistic conditions.
-We feel that there is potential to significantly improve the performance of our current approach.
-For example, we currently buffer all results at each server before sending a reply.
-Streaming results back to the client as they are received may result in higher link utilization and increased efficiency.
+While our results are promising, we would like to take a more principled
+approach to determine the theoretical gains possible for applying this
+technique in order to clearly identify potential bottlenecks.  For example, we
+have implemented a more realistic datacenter like topology with potentially
+large number of servers. However, we find that our system requires a lot fine
+tuning of minimum bandwidth guarantees for more complex systems. Theoretical
+understanding of our system and approach could potentially enable us to
+automate this fine-tuning process, which in turn allows us to perform a more
+thorough evaluation on a larger scale with more realistic conditions. We feel
+that there is potential to significantly improve the performance of our current
+approach.  For example, we currently buffer all results at each server before
+sending a reply.  Streaming results back to the client as they are received may
+result in higher link utilization and increased efficiency.
 
-There are many opportunities to combine this work with appropriate key placement strategies for further performance optimizations.
-For heterogeneous networks, a key placement strategy which is network and workload aware may provide significant performance advantages.
-Our current query execution model is very simplistic and could be expanded to support richer queries and distributed index data structures.
-We would also like to investigate other query types such as aggregation which we suspect would also benefit from our approach.
-This presents new challenges in network scheduling, especially in the face of concurrent queries of multiple different types.
+There are many opportunities to combine this work with appropriate key
+placement strategies for further performance optimizations.  For heterogeneous
+networks, a key placement strategy which is network and workload aware may
+provide significant performance advantages. On the flip side of data placement
+problem is the data shipment problem where we can exploit our system's network
+awareness to decide which node the data should be shipped to in order to
+perform efficient join operation. Our current query execution model is very
+simplistic and could be expanded to support richer queries and distributed
+index data structures.  We would also like to investigate other query types
+such as aggregation which we suspect would also benefit from our approach.
+This presents new challenges in network scheduling, especially in the face of
+concurrent queries of multiple different types.
 
-Our current system also does not deal with the possibility of updates or replication.
-There is potential to further exploit SDN techniques by allowing low latency updates to data in the presence of the high bandwidth flows used for query processing.
-Replication provides further opportunity for optimization as flows can be routed to multiple possible locations to avoid bottleneck links.
+Another promising direction is to consider the other kinds of network
+topologies, such as fat tree topologies, where there exists multiple path
+between any two hosts in the network. This enables our controller to use a
+combination of network aware routing and minimum bandwidth guarantee techniques
+to improve the overall performance of each query. We foresee that we will
+evaluate our network-aware system against regular routing protocols such as
+Open Shortest Path First routing to study the networking behaviours..
+
+Our current system also does not deal with the possibility of updates or
+replication.  There is potential to further exploit SDN techniques by allowing
+low latency updates to data in the presence of the high bandwidth flows used
+for query processing.  Replication provides further opportunity for
+optimization as flows can be routed to multiple possible locations to avoid
+bottleneck links or hosts.
+
+We need to thoroughly study each of the aforementioned techniques to fully
+understand their pros and cons in order to integrate all of them as a whole to
+provide the optimal join operation for distributed NoSQL databases.
 
@@
 
-In this work, we examined opportunties to apply software-defined network techniques to provide more efficient query processing over NoSQL databases.
-We demonstrated a simple technique using smart network scheduling to improve the performance of distributed joins.
-Our technique provides a significant performance improvement over uninformed switching techniques.
+In this work, we examined opportunities to apply software-defined network
+techniques to provide more efficient query processing over NoSQL databases.  We
+demonstrated a simple technique using smart network scheduling to improve the
+performance of distributed joins.  Our preliminary technique provides a
+significant performance improvement over uninformed switching techniques.
 
-We also presented several opportunties to further exploit of software-defined network techniques to improve query processing.
-As software-defined networking becomes increasingly common in datacentres, we feel similar techniques will become increasingly prevalent.
-Little work has been done to exploit software-defined networking in the context of NoSQL databases and distributed data management.
+We also presented several opportunities to further exploit of software-defined
+network techniques to improve query processing.  As software-defined networking
+becomes increasingly common in datacenters, we feel similar techniques will
+become increasingly prevalent.  Little work has been done to exploit
+software-defined networking in the context of NoSQL databases and distributed
+data management.
 

### 2014-08-08 paper/paper.tex
Add more details on controller implementation

@@
 \usepackage{microtype}
+\usepackage{listings}
 
@@
 However, we configure several queues for each switch port which allow us to provide bandwidth reservations to support variation in expected traffic volume.
+Once we have learned the port mapping, we simply install a rule that also matches on the DSCP bits and assigns the packet to an appropriate queue as in Figure~\ref{fig:scheduling}, which shows some sample code from our controller.
+
+\begin{figure*}
+    \begin{lstlisting}[language=Python]
+        # Only perform queue assignment for IP packets
+        if(packet.type == 0x0800):
+          # Extract the DSCP bits from the ToS (ignoring ECN)
+          queue_num = (packet.payload.tos >> 2)
+
+          # Send the packet to the appropriate queue
+          msg.actions.append(of.ofp_action_enqueue(port=port, queue_id=queue_num))
+        else:
+          # Send all non-IP packets to a queue with higher priority
+          msg.actions.append(of.ofp_action_enqueue(port=port, queue_id=0))
+    \end{lstlisting}
+    \caption{Simple flow scheduling to queues}\label{fig:scheduling}
+\end{figure*}
 
@@
 We use Open vSwitch~\cite{Pfaff2009} as our switch which supports queues with varying bandwidth guarantees.
-Our controller uses the POX~\cite{Gude2008} platform and OpenFlow~\cite{McKeown2008} 1.0.
+Queuing is performed by the hierarchical token bucket queuing discipline in the Linux kernel.
+We simply specify a set of queues for each switch port with varying minimum bandwidth guarantees.
+Our controller is written in Python using the POX~\cite{Gude2008} platform and OpenFlow~\cite{McKeown2008} 1.0.
 

### 2014-08-08 paper/paper.tex
Minor report edits

@@
 
-There are other advanced join operations which may benefit from software-defined
-networking. Ntarmos et al.~\cite{Ntarmos2014} uses novel indexing techniques to
-improve the performance of top-k joins for NoSQL databases which in turn
-reduces both network latency and network traffics. This work did not explicit
-use software-defined networking, however the amount of network traffic is one
-important metric in their evaluation which hints possible network related
-work in this domain.
+There are other advanced join operations which may benefit from software-defined networking.
+Ntarmos et al.~\cite{Ntarmos2014} uses novel indexing techniques to improve the performance of top-k joins for NoSQL databases which in turn reduces both network latency and traffic.
+This work did not explicitly use software-defined networking.
+However, the amount of network traffic is a key metric in their evaluation, hinting at the possibility to successfully apply SDN techniques.
 
@@
 
-While our results are promising, we would like to take a more principled
-approach to determine the theoretical gains possible for applying this
-technique in order to clearly identify potential bottlenecks.  For example, we
-have implemented a more realistic datacenter like topology with potentially
-large number of servers. However, we find that our system requires a lot fine
-tuning of minimum bandwidth guarantees for more complex systems. Theoretical
-understanding of our system and approach could potentially enable us to
-automate this fine-tuning process, which in turn allows us to perform a more
-thorough evaluation on a larger scale with more realistic conditions. We feel
-that there is potential to significantly improve the performance of our current
-approach.  For example, we currently buffer all results at each server before
-sending a reply.  Streaming results back to the client as they are received may
-result in higher link utilization and increased efficiency.
+While our results are promising, we would like to take a more principled approach to determine the theoretical gains possible for applying this
+technique in order to clearly identify potential bottlenecks.
+For example, we have implemented a more realistic datacenter like topology with potentially large number of servers.
+However, we found that our system requires a lot of fine tuning for more complex topologies.
+A more thorough theoretical understanding of our system and approach could potentially enable us to automate this process, which in turn would allows us a more thorough evaluation on a larger scale with more realistic conditions.
+We feel that there is potential to significantly improve the performance of our current approach in several ways.
+For example, we currently buffer all results at each server before sending a reply.
+Streaming results back to the client as they are received may result in higher link utilization and increased efficiency.
 
-There are many opportunities to combine this work with appropriate key
-placement strategies for further performance optimizations.  For heterogeneous
-networks, a key placement strategy which is network and workload aware may
-provide significant performance advantages. On the flip side of data placement
-problem is the data shipment problem where we can exploit our system's network
-awareness to decide which node the data should be shipped to in order to
-perform efficient join operation. Our current query execution model is very
-simplistic and could be expanded to support richer queries and distributed
-index data structures.  We would also like to investigate other query types
-such as aggregation which we suspect would also benefit from our approach.
-This presents new challenges in network scheduling, especially in the face of
-concurrent queries of multiple different types.
+There are many opportunities to combine this work with appropriate key placement strategies for further performance optimizations.
+A key placement strategy which is network and workload aware may provide significant performance advantages.
+On the flip side of the data placement problem is the problem of data shipping, where we can exploit our system's network awareness to decide where data should be shipped to execute joins most efficiently.
+Futhermore, our current query execution model is very simplistic and could be expanded to support richer queries and distributed index data structures.
+We would also like to investigate other query types such as aggregation which we suspect would also benefit from our approach.
+This presents new challenges in network scheduling, especially in the face of concurrent queries of multiple different types.
 
-Another promising direction is to consider the other kinds of network
-topologies, such as fat tree topologies, where there exists multiple path
-between any two hosts in the network. This enables our controller to use a
-combination of network aware routing and minimum bandwidth guarantee techniques
-to improve the overall performance of each query. We foresee that we will
-evaluate our network-aware system against regular routing protocols such as
-Open Shortest Path First routing to study the networking behaviours..
+Another promising direction is to consider the other network topologies, such as fat trees, where there exist multiple paths between any two hosts.
+This provide opportunities to use a combination of network aware routing and minimum bandwidth guarantees to improve the overall performance of each query.
+We foresee that we will evaluate our network-aware system against regular routing protocols such as Open Shortest Path First (OSPF).
 
-Our current system also does not deal with the possibility of updates or
-replication.  There is potential to further exploit SDN techniques by allowing
-low latency updates to data in the presence of the high bandwidth flows used
-for query processing.  Replication provides further opportunity for
-optimization as flows can be routed to multiple possible locations to avoid
-bottleneck links or hosts.
-
-We need to thoroughly study each of the aforementioned techniques to fully
-understand their pros and cons in order to integrate all of them as a whole to
-provide the optimal join operation for distributed NoSQL databases.
+Our current system also does not deal with the possibility of updates or replication.
+There is potential to further exploit SDN techniques by allowing low latency updates to data in the presence of the high bandwidth flows used
+for query processing.
+Replication also provides further opportunities for optimization as flows can be routed to multiple possible locations to avoid bottleneck links or hosts.
 
@@
 
-In this work, we examined opportunities to apply software-defined network
-techniques to provide more efficient query processing over NoSQL databases.  We
-demonstrated a simple technique using smart network scheduling to improve the
-performance of distributed joins.  Our preliminary technique provides a
-significant performance improvement over uninformed switching techniques.
+In this work, we examined opportunities to apply software-defined network techniques to provide more efficient query processing over NoSQL databases.
+We demonstrated a simple technique using smart network scheduling to improve the performance of distributed joins.
+Our preliminary technique provides a significant performance improvement over uninformed switching techniques.
 
-We also presented several opportunities to further exploit of software-defined
-network techniques to improve query processing.  As software-defined networking
-becomes increasingly common in datacenters, we feel similar techniques will
-become increasingly prevalent.  Little work has been done to exploit
-software-defined networking in the context of NoSQL databases and distributed
-data management.
+We also presented several opportunities to further exploit of software-defined network techniques to improve query processing.
+As software-defined networking becomes increasingly common in datacenters, we feel similar techniques will become increasingly prevalent.
+Little work has been done to exploit software-defined networking in the context of NoSQL databases and distributed data management.
 

### 2014-08-08 plot_src/simple_dc_varying_bandwidth.py
Adds the final figure with varying link capacity

@@ -1,0 +1,102 @@
+#!/usr/bin/env python
+# a stacked bar plot with errorbars
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+import re
+from os.path import expanduser
+import confidence_interval_calculator as CI
+import save_fig as saver
+
+
+def get_overall_arr(key_word, data_dir):
+    rtn = []
+    for i in range(1,6):
+        data = open(data_dir + 'client_' + (str(i)) + '.out' , "r")
+        for line in data:
+            if(re.search(key_word, line)):
+                tokens = line.split(" ")
+                rtn.append(float(tokens[len(tokens) - 3]))
+    return rtn
+
+PATTERN = [ "/" , "\\" ,"o", "*", "O", "x", ".", '-', "|" , "+"]
+base_dir = '/mininet/src/result/various_link_bandwidth/'
+data_dirs = [base_dir + '0_8mbps/',
+             base_dir + '0_9mbps/',
+             base_dir + '1_0mbps/',
+             base_dir + '1_1mbps/',
+             base_dir + '1_2mbps/'
+         ]
+
+
+base_mean = []
+base_err = []
+
+# return values will be filled into rtn_mean, rtn_std
+def get_results(key_word):
+    for i in range(len(data_dirs)):
+        overall_arr = get_overall_arr(key_word, data_dirs[i])
+        stdmean_stderr = CI.get_conf_int(overall_arr)
+        base_mean.append(int(stdmean_stderr[0] / 100) / 10.0)
+        base_err.append(int(stdmean_stderr[1] / 100) / 10.0)
+
+
+get_results('Total time')
+
+comp_mean = []
+comp_err = []
+
+base_dir = '/mininet/src/result/various_link_bandwidth_scheduling/'
+data_dirs = [base_dir + '0_8mbps/',
+             base_dir + '0_9mbps/',
+             base_dir + '1_0mbps/',
+             base_dir + '1_1mbps/',
+             base_dir + '1_2mbps/'
+         ]
+# return values will be filled into rtn_mean, rtn_std
+def get_results_new(key_word):
+    for i in range(len(data_dirs)):
+        overall_arr = get_overall_arr(key_word, data_dirs[i])
+        stdmean_stderr = CI.get_conf_int(overall_arr)
+        comp_mean.append(int(stdmean_stderr[0] / 100) / 10.0)
[…]

### 2014-08-08 paper/paper.tex
Switch away from ACM alternate style
Use full size biliography
Adds last figure & last bit text of the paper

@@
-\documentclass{sig-alternate-2013}
+\documentclass{acm_proc_article-sp}
 
@@
     \centering
-    \includegraphics[scale=2.5]{figures/Join_micro_benchmark.png}
+    \includegraphics[scale=2]{figures/Join_micro_benchmark.png}
     \caption{Micro benchmark setup}\label{fig:micro_benchmark}
@@
 
+To understand how link capacity affects the perform.
+We also tested our micro benchmark by varying the link capacity.
+The result is shown in figure~\ref{fig:varying_bd}.
+During each of the experiment, the system parameters, e.g.
+the minimum bandwidth guarantee for each server (server 1 will always have min-guarantee off 80\% of the total link capacity), are fixed and we only varied the link capacity.
+We see that our system can adaptively work with different link capacities and our system is uniformly performing better than the baseline system without scheduling.
+The next promising step is to increase link capacity at larger steps and study how system will react and try to learn if we need to adjust system adaptivity to the link capacity.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=0.47]{figures/simple_topo_varying_bd.png}
+    \caption{Query completion time with varying link capacity}\label{fig:varying_bd}
+\end{figure}
+
 \section{Future Work}
@@
 On the flip side of the data placement problem is the problem of data shipping, where we can exploit our system's network awareness to decide where data should be shipped to execute joins most efficiently.
-Futhermore, our current query execution model is very simplistic and could be expanded to support richer queries and distributed index data structures.
+Furthermore, our current query execution model is very simplistic and could be expanded to support richer queries and distributed index data structures.
 We would also like to investigate other query types such as aggregation which we suspect would also benefit from our approach.
@@
 
-\let\theOLDbibliography\thebibliography\renewcommand{\thebibliography}[1]{\theOLDbibliography{#1}%
-\item[]\vspace*{0.5mm}}
-
 \bibliographystyle{abbrv}
-{\scriptsize
 \bibliography{paper}
-}
 

### 2014-08-08 paper/paper.tex
Minor paper edits

@@
 We avoid these bits to allow ECN to operate uninhibited.
+Our system do not currently make use of ECN, but this could be worth exploring in the future.
 The remaining six bits are the differentiated services code point (DSCP) which are used to provide varying QoS guarantees for different flows.
@@
 This is expected due to the data distribution we defined above.
-We can also see that when we enabled our network traffic scheduling mechanisms, the completion time for server 3 and server 4 slowed down slightly.
-This is because most of the bandwidth will be allocated to help server 1 to finish its task as soon as possible.
-However, The end result is that traffic scheduling reduces the average subquery completion time, which in turn reduces the overall query execution time.
+We can also see that when we enabled our network traffic scheduling mechanisms, the completion time for servers 3 and 4 increased slightly.
+This is because most of the bandwidth is allocated to help server 1 to finish its task as soon as possible.
+However, the end result is that traffic scheduling reduces the average subquery completion time, which in turn reduces the overall query execution time.
 We see that our experimental results match with our expectations.
@@
 We also tested our micro benchmark by varying the link capacity.
-The result is shown in figure~\ref{fig:varying_bd}.
-During each of the experiment, the system parameters, e.g.
-the minimum bandwidth guarantee for each server (server 1 will always have min-guarantee off 80\% of the total link capacity), are fixed and we only varied the link capacity.
-We see that our system can adaptively work with different link capacities and our system is uniformly performing better than the baseline system without scheduling.
-The next promising step is to increase link capacity at larger steps and study how system will react and try to learn if we need to adjust system adaptivity to the link capacity.
+The result is shown in Figure~\ref{fig:varying_bd}.
+During each of the experiments, the system parameters, e.g.\ the minimum bandwidth guarantees for each server, are fixed while varying the total capacity of each link.
+We see that our system can adaptively work with different link capacities and our system consistently outperforms the baseline with no traffic scheduling.
 
 \begin{figure}
-    \centering
     \includegraphics[scale=0.47]{figures/simple_topo_varying_bd.png}
@@
 technique in order to clearly identify potential bottlenecks.
-For example, we have implemented a more realistic datacenter like topology with potentially large number of servers.
+For example, we have implemented a more realistic data centre like topology with potentially large number of servers.
 However, we found that our system requires a lot of fine tuning for more complex topologies.
@@
 We also presented several opportunities to further exploit of software-defined network techniques to improve query processing.
-As software-defined networking becomes increasingly common in datacenters, we feel similar techniques will become increasingly prevalent.
+As software-defined networking becomes increasingly common in data centres, we feel similar techniques will become increasingly prevalent.
 Little work has been done to exploit software-defined networking in the context of NoSQL databases and distributed data management.

### 2014-08-08 paper/paper.tex
Convert graphs to PDF

@@
     \centering
-    \includegraphics[scale=0.47]{figures/simple_topo_overall.png}
+    \includegraphics[scale=0.47]{figures/simple_topo_overall.pdf}
     \caption{Query completion time}\label{fig:micro_benchmark_total_time}
@@
     \centering
-    \includegraphics[scale=0.47]{figures/simple_topo_closer_8.png}
+    \includegraphics[scale=0.47]{figures/simple_topo_closer_8.pdf}
     \caption{Sub query completion time}\label{fig:sub_query_completion_8}
@@
 \begin{figure}
-    \includegraphics[scale=0.47]{figures/simple_topo_varying_bd.png}
+    \includegraphics[scale=0.47]{figures/simple_topo_varying_bd.pdf}
     \caption{Query completion time with varying link capacity}\label{fig:varying_bd}

### 2014-08-08 plot_src/save_fig.py

@@ -6,7 +6,7 @@
 import os
 import sys
 
-def save(user_plt, path, ext='png', close=True, verbose=True):
+def save(user_plt, path, ext='pdf', close=True, verbose=True):
     """Save a figure from pyplot.
 
     Parameters

### 2014-08-08 plot_src/simple_dc_closer_8.py

@@ -20,7 +20,7 @@ def get_overall_arr(key_word, data_dir):
     return rtn
 
 PATTERN = [ "/" , "\\" ,"o", "*", "O", "x", ".", '-', "|" , "+"]
-base_dir = '/mininet/src/result/1mbps/'
+base_dir = '/home/mmior/apps/kvtest/src/result/1mbps/'
 data_dirs = [base_dir + 'norm/',
              base_dir + 'scheduling_queue_8_for_h1/']
 
@@ -82,8 +82,8 @@ plt.xlabel('Server hosts')
 plt.xticks(ind+width, ('h4', 'h3', 'h2', 'h1') )
 # plt.yticks(np.arange(0,46,5))
 plt.legend((p0[0], p1[0]),
-           ('Without Scheduling',
-            'With Scheduling'))
+           ('Without scheduling',
+            'With scheduling'))
            # loc = 'upper left', prop={'size':12},
            # ncol is # of columns for the legend
            # ncol=2)

### 2014-08-08 plot_src/simple_dc_overall.py

@@ -20,7 +20,7 @@ def get_overall_arr(key_word, data_dir):
     return rtn
 
 PATTERN = [ "/" , "\\" ,"o", "*", "O", "x", ".", '-', "|" , "+"]
-base_dir = '/mininet/src/result/1mbps/'
+base_dir = '/home/mmior/apps/kvtest/src/result/1mbps/'
 data_dirs = [base_dir + 'norm/',
              base_dir + 'scheduling_queue_5_for_h1/',
              base_dir + 'scheduling_queue_6_for_h1/',
@@ -85,8 +85,8 @@ plt.xlabel('Minimum bandwidth guarantee for the largest flow in Mbps')
 plt.xticks(ind+width, ('0.5', '0.6', '0.7', '0.8', '0.9', '1.0') )
 # plt.yticks(np.arange(0,46,5))
 plt.legend((p0[0], p1[0]),
-           ('Without Scheduling',
-            'With Scheduling'))
+           ('Without scheduling',
+            'With scheduling'))
            # loc = 'upper left', prop={'size':12},
            # ncol is # of columns for the legend
            # ncol=2)

### 2014-08-08 plot_src/simple_dc_varying_bandwidth.py

@@ -20,7 +20,7 @@ def get_overall_arr(key_word, data_dir):
     return rtn
 
 PATTERN = [ "/" , "\\" ,"o", "*", "O", "x", ".", '-', "|" , "+"]
-base_dir = '/mininet/src/result/various_link_bandwidth/'
+base_dir = '/home/mmior/apps/kvtest/src/result/various_link_bandwidth/'
 data_dirs = [base_dir + '0_8mbps/',
              base_dir + '0_9mbps/',
              base_dir + '1_0mbps/',
@@ -46,7 +46,7 @@ get_results('Total time')
 comp_mean = []
 comp_err = []
 
-base_dir = '/mininet/src/result/various_link_bandwidth_scheduling/'
+base_dir = '/home/mmior/apps/kvtest/src/result/various_link_bandwidth_scheduling/'
 data_dirs = [base_dir + '0_8mbps/',
              base_dir + '0_9mbps/',
              base_dir + '1_0mbps/',
@@ -90,7 +90,7 @@ plt.ylabel('Query completion time in seconds')
 plt.xlabel('Link capacity in Mbps')
 plt.xticks((0,1,2,3,4), ('0.8', '0.9', '1.0', '1.1', '1.2') )
 plt.legend(loc='upper right')
-# plt.legend([base, comp_line], ['Without Scheduling', 'With Scheduling'])
+# plt.legend([base, comp_line], ['Without scheduling', 'With scheduling'])
 
 # plt.yticks(np.arange(0,46,5))
 # plt.legend((p0[0], p1[0]),

### 2014-08-08 paper/paper.tex
Various paper corrections

@@
 
-This results in the issue of stragglers we presented previously.
+This results in the issue of stragglers which we presented previously.
 Note that the common technique of issuing a request multiple times to avoid stragglers is not helpful in our scenario.
-As previously stated, our performance is bottlenecked by the available bandwidth.
+As previously stated, the performance of our system is bottlenecked by the available bandwidth.
 Stragglers exist in our system because of a lack of available bandwidth as opposed to variations in node performance (which is what multiple issuance is designed to mitigate).
@@
 
-First, we want to demonstrate the proof-of-concept of our design by evaluating our system with the simple topology shown in Figure~\ref{fig:micro_benchmark}.
+First, we demonstrate a simple proof-of-concept of our design by evaluating our system with the simple topology shown in Figure~\ref{fig:micro_benchmark}.
 We engineered the client query such that each server holds a different portion of the relevant data for the join operation.
@@
 The average completion time without scheduling for each client query is over 180 seconds and also shows significant variation.
-To provide predictable performance with our scheduling mechanism, we must first understand the characteristics of the costs for each component of the join operation.
-In our case it is easy to predict the costs because we assigned different portions of relevant data to each server.
-However, it is also easy in a real deployment scenario as the data distribution among servers is typically available or can be easily computed.
+To provide predictable performance with our scheduling mechanism, we must first understand the costs for each component of the join operation.
+In our case, it is easy to predict the costs because we assigned different portions of relevant data to each server.
+However, it is also straightforward in a real deployment scenario as the data distribution among servers is typically available or can be easily computed.
 Recall that by design, server 1 is responsible for transmitting the largest flows between other servers and the client.
@@
 
-Though we want server 1 to transfer its data as fast as possible, we do not want to starve other servers.
+Though we want server 1 to transfer its data as fast as possible, we do not want to starve other servers of bandwidth.
 Therefore, we must also have reasonable minimum bandwidth guarantees for other servers.
-Through experimentation, we found the most suitable minimum bandwidth guarantees on the 1 Mbps link for servers 2, 3, and 4 are 0.3 Mbps, 0.2 Mbps and 0.1 Mbps respectively.
+Through experimentation, we found the most suitable minimum bandwidth guarantees on the 1 Mbps link for servers 2, 3, and 4 are 0.3 Mbps, 0.2 Mbps, and 0.1 Mbps respectively.
 We fixed these parameters and varied the minimum bandwidth guarantee for server 1 to study the network behaviour.
@@
 This is caused by a significant increase in the completion time for other flows, hurting the overall performance.
-In the best case, with 0.8 Mbps minimum bandwidth guarantee for server 1, the average completion time of the client query is 154 seconds which gives us almost 15\% performance gain over the baseline implementation.
+In the best case, with a 0.8 Mbps minimum bandwidth guarantee for server 1, the average completion time of the client query is 154 seconds which gives us almost a 15\% performance gain over the baseline implementation.
 
@@
 
-To understand how link capacity affects the perform.
-We also tested our micro benchmark by varying the link capacity.
-The result is shown in Figure~\ref{fig:varying_bd}.
+We also examined the effects of varying link capacity on our systems, with results shown in Figure~\ref{fig:varying_bd}.
 During each of the experiments, the system parameters, e.g.\ the minimum bandwidth guarantees for each server, are fixed while varying the total capacity of each link.
@@
 technique in order to clearly identify potential bottlenecks.
-For example, we have implemented a more realistic data centre like topology with potentially large number of servers.
+We have tested our system under a more realistic data centre topology with a larger number of servers.
 However, we found that our system requires a lot of fine tuning for more complex topologies.
-A more thorough theoretical understanding of our system and approach could potentially enable us to automate this process, which in turn would allows us a more thorough evaluation on a larger scale with more realistic conditions.
-We feel that there is potential to significantly improve the performance of our current approach in several ways.
+A better theoretical understanding of our system and approach will enable us to automate this process, which in turn would allow us to perform a more thorough evaluation on a larger scale and with more realistic conditions.
+We believe that there is potential to significantly improve the performance of our current approach in several ways.
 For example, we currently buffer all results at each server before sending a reply.
@@
 
-Another promising direction is to consider the other network topologies, such as fat trees, where there exist multiple paths between any two hosts.
-This provide opportunities to use a combination of network aware routing and minimum bandwidth guarantees to improve the overall performance of each query.
-We foresee that we will evaluate our network-aware system against regular routing protocols such as Open Shortest Path First (OSPF).
+Another promising direction is to consider other network topologies, such as fat trees, where there exist multiple paths between any two hosts.
+This provides opportunities to use a combination of network-aware routing and minimum bandwidth guarantees to improve the overall performance of each query.
+We can also exploit the multiple available paths to attempt to avoid flow collisions.
+We foresee that we will evaluate our network-aware system against standard routing protocols such as Open Shortest Path First (OSPF).
 
@@
 
-In this work, we examined opportunities to apply software-defined network techniques to provide more efficient query processing over NoSQL databases.
+In this work, we examined methods of applying software-defined network techniques to provide more efficient query processing over NoSQL databases.
 We demonstrated a simple technique using smart network scheduling to improve the performance of distributed joins.
@@
 
-We also presented several opportunities to further exploit of software-defined network techniques to improve query processing.
-As software-defined networking becomes increasingly common in data centres, we feel similar techniques will become increasingly prevalent.
+We also presented additional opportunities to further exploit software-defined networking techniques to improve query processing.
 Little work has been done to exploit software-defined networking in the context of NoSQL databases and distributed data management.
+As software-defined networking becomes increasingly common in data centres, we believe similar techniques will become increasingly prevalent.
 
